### 2023-08-30 Paper/main.tex
update

% !TEX encoding = UTF-8
% !TEX program = pdflatex
% !TeX spellcheck = en_GB
% !BIB = biber

\documentclass[english]{article}
\usepackage{babel}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[obeyspaces]{url}
\graphicspath{{./images/}}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true, 
	linkcolor=blue, 
	filecolor=blue, 
	citecolor = black,       
	urlcolor=blue, 
}
\usepackage{listings}
\lstset{
  xleftmargin=15pt,
  xrightmargin=0pt,
  framexleftmargin=0pt,
  framexrightmargin=0pt,
  basicstyle={\fontsize{9pt}{10pt}\ttfamily},
  columns=flexible,
  numbers=left,
  numbersep=10pt,
  numberstyle={\fontsize{9pt}{11pt}\selectfont\color[rgb]{0.4,0.4,0.1}},
  keepspaces=true,
  showstringspaces=false,
  identifierstyle=\color[rgb]{0.1,0.1,0.1},
  keywordstyle=\color{blue},
  commentstyle=\color[rgb]{0,0.3,0},
  morekeywords={rule, lemma},
  morekeywords=[2]{let, in},
  morekeywords=[3]{Fr, pk},
  morekeywords=[4]{In, Out},
  morekeywords=[5]{senc, aenc, h}
  morecomment=[s][keywordstyle3]{/*}{/},
  keywordstyle=\color[rgb]{0.44,0.57,0.65},
  stringstyle=\color{green},
  keywordstyle=[2]{\color[rgb]{0.86,0.57,0.18}},
  keywordstyle=[3]{\bfseries\color[rgb]{0,0.3,0.2}}
  }
\usepackage{biblatex}
\addbibresource{thud.bib}

\title{LTL+P Based Framework for Intrusion Detection}
\author{Zanolin Lorenzo}

\begin{document}

\maketitle

\begin{abstract}
The purpose of this paper is to introduce MONID, which is a framework created for system intrusion detection. This framework uses EAGLE, a rich and effectively monitorable logic, to express intrusion patterns using temporal logic formulas; EAGLE's ability to include data values and parameterized recursive equations makes it possible to represent security threats that include complex temporal event sequences and attacks with intrinsically statistical signatures succinctly. This tool can be used in offline and real-time scenarios. The implementation uses an algorithm for online monitoring that matches descriptions of the lack of an assault with indications of system execution; an alarm is set off whenever the standard is broken.
\end{abstract}

\tableofcontents
\newpage

\section{Introduction}

\printbibliography

\end{document}

### 2023-08-30 Paper/main.tex
update

@@
 \section{Introduction}
+Even with all of the advances in computer security research, totally safe computer systems remain a long way off.
+Almost every large and complicated computer system nowadays contains vulnerabilities.
+Intrusion detection means maintaining constant surveillance on a system in order to detect any misuse of these weak areas as soon as feasible so that they can be repaired.
+
+There are three Intrusion Detection System approaches, according to the literature: \textit{signature-based}\cite{gao2014cyber}, \textit{anomaly-based}\cite{jyothsna2011review} and \textit{hybrid}.
+The first approach aims to identify patterns and match them with known signs of intrusions relying on a database of previous intrusions.
+If activity within the network matches the “signature” of an attack or breach from the database, the detection system creates an alert.
+This approach has a low false-alarm rate, but it requires us to know the patterns of security attacks in advance and previously unknown attacks would go undetected.
+In contrast, \textit{anomaly-based} is capable to detect new attacks since an alarm is raised if an observed behavior deviates significantly from pre-learned normal behavior.
+Finally, a \textit{hybrid system} combines the best of both worlds by looking at patterns and one-off events, a Hybrid Intrusion Detection system can flag new and existing intrusion strategies.
 

### 2023-08-30 Paper/main.tex
update

@@
 
-There are three Intrusion Detection System approaches, according to the literature: \textit{signature-based}\cite{gao2014cyber}, \textit{anomaly-based}\cite{jyothsna2011review} and \textit{hybrid}.
+There are three Intrusion Detection System approaches, according to the literature: \textit{signature-based}\cite{gao2014cyber}, \textit{anomaly-based}\cite{jyothsna2011review} and \textit{hybrid}\cite{garg2016hybrid}.
 The first approach aims to identify patterns and match them with known signs of intrusions relying on a database of previous intrusions.
@@
 
+In this paper we will focus on \textit{signature-based} approach using temporal logic; we use EAGLE\cite{barringer2004rule,barringer2003eagle} to specify a system's attack-safe behavior.
+EAGLE allows recursively built temporal formulas, parameterizable by both logical formulas and data expressions, across three primitive modalities: "next", "previous", and "concatenation".
+The logic allows us to describe temporal patterns involving reasoning about data-values observed in individual events, allowing us to design attacks with fundamentally statistical characteristics.
+Password guessing attacks and ICMP-flood denial of service attacks are two examples.
+
 \printbibliography

### 2023-08-30 Paper/main.tex
update

@@
 EAGLE allows recursively built temporal formulas, parameterizable by both logical formulas and data expressions, across three primitive modalities: "next", "previous", and "concatenation".
-The logic allows us to describe temporal patterns involving reasoning about data-values observed in individual events, allowing us to design attacks with fundamentally statistical characteristics.
-Password guessing attacks and ICMP-flood denial of service attacks are two examples.
+The logic allows us to describe temporal patterns involving reasoning about data-values observed in individual events.
+Unlike LTL, EAGLE allowa us to design attacks with fundamentally statistical characteristics; password guessing attacks and ICMP-flood denial of service attacks are two examples.
+We'll see how the implementation, MONID, examines the event stream to see if the monitored formula (which is used to hold the pertinent summary of the system) is being broken.
+
+%%%%%   DA RISCRIVERE ALLA FINE
+In Section 2 we will analyze the basics of EAGLE.
+In Section 3 we will analyze MONID framework and some evalutations on the performances.
+In Section 4 we will see some attacks specifications.
+In section 5 we will conclude.
+%%%%5
 

### 2023-08-30 Paper/main.tex
update

@@
 \documentclass[english]{article}
+\usepackage{amsmath}
+\usepackage{wasysym}
 \usepackage{babel}
@@
 
-\title{LTL+P Based Framework for Intrusion Detection}
+\title{A Temporal Logic Based Framework for Intrusion Detection}
 \author{Zanolin Lorenzo}
@@
 
+\section{EAGLE, a Temporal Monitoring Logic}
+According to~\cite{barringer2004program}, EAGLE offers a succinct but powerful set of primitives, essentially supporting recursive parameterized equations, with a minimal/maximal fix-point semantics together with three temporal operators: next-time ($\ocircle$), previous-time ($\odot$), and concatenation ($\cdot$).
+\subsection{Basics and syntax}
+In EAGLE recursion definitions are supported; in the current framework we can define the following definitions:
+
+\begin{align*}
+& \underline{\text{max}}\ \text{Always}(\underline{\text{Form}}\ F) = F \land \bigcirc \text{Always}(F) \\
+& \underline{\text{min}}\ \text{Sometime}(\underline{\text{Form}}\ F) = F \lor \bigcirc \text{Sometime}(F) \\
+\end{align*}
+
+We will use it to create formulas that describe the absence of some attacks and given a trace of events $\sigma$ generated from the logs we will check whether
+
 \printbibliography

### 2023-08-30 Paper/main.tex
update

@@
 In EAGLE recursion definitions are supported; in the current framework we can define the following definitions:
-
 \begin{align*}
-& \underline{\text{max}}\ \text{Always}(\underline{\text{Form}}\ F) = F \land \bigcirc \text{Always}(F) \\
+\underline{\text{max}}\ \text{Always}(\underline{\text{Form}}\ F) = F \land \bigcirc \text{Always}(F) \\
 & \underline{\text{min}}\ \text{Sometime}(\underline{\text{Form}}\ F) = F \lor \bigcirc \text{Sometime}(F) \\

### 2023-08-31 Paper/main.tex
update

--- 2023-08-30 Paper/main.tex
@@
 \subsection{Basics and syntax}
-In EAGLE recursion definitions are supported; in the current framework we can define the following definitions:
-
+In EAGLE recursion definitions are supported; in the current framework we can build the following definitions:
 \begin{align*}
@@
 \end{align*}
+
+As we can see, rules are parameterized, which means that we will be able to write EAGLE formulas such as $\text{Always}(\text{Sometime}(x>0))$.
+Also, the $\text{Always}$ operator is defined as maximal solution of the equation $X=F\land \ocircle X$ and the $\text{Sometime}$ operator represents the minimal solution to the equation $X=F\lor \ocircle X$.
 

### 2023-08-31 Paper/main.tex
update

--- 2023-08-30 Paper/main.tex
@@
 \usepackage{wasysym}
+\usepackage{amssymb}
 \usepackage{babel}
@@
 \section{EAGLE, a Temporal Monitoring Logic}
-According to~\cite{barringer2004program}, EAGLE offers a succinct but powerful set of primitives, essentially supporting recursive parameterized equations, with a minimal/maximal fix-point semantics together with three temporal operators: next-time ($\ocircle$), previous-time ($\odot$), and concatenation ($\cdot$).
-\subsection{Basics and syntax}
-In EAGLE recursion definitions are supported; in the current framework we can define the following definitions:
+According to~\cite{barringer2004program}, EAGLE offers a succinct but powerful set of primitives, essentially supporting recursive parameterized equations, with a minimal/maximal fix-point semantics together with three temporal operators: next-time ($\bigcirc$), previous-time ($\odot$), and concatenation ($\cdot$).
+\subsection{Basics}
+In EAGLE recursion definitions are supported; in the current framework we can build the following definitions:
 \begin{align*}
-\underline{\text{max}}\ \text{Always}(\underline{\text{Form}}\ F) = F \land \bigcirc \text{Always}(F) \\
-& \underline{\text{min}}\ \text{Sometime}(\underline{\text{Form}}\ F) = F \lor \bigcirc \text{Sometime}(F) \\
+& \underline{\text{max}}\ \text{Always}(\underline{\text{Form}}\ F) = F \land \bigcirc \text{Always}(F) \\
+& \underline{\text{min}}\ \text{Eventually}(\underline{\text{Form}}\ F) = F \lor \bigcirc \text{Eventually}(F) 
+\end{align*}
+
+As we can see, rules are parameterized by an EAGLE formula (of type $\underline{\text{Form}}$), which means that we will be able to write EAGLE formulas such as $\text{Always}(\text{Eventually}(x>0))$.
+Also, the $\text{Always}$ operator is defined as maximal solution of the equation $X=F\land \bigcirc X$ and the $\text{Sometime}$ operator represents the minimal solution to the equation $X=F\lor \bigcirc X$.
+%%  QUI MANCA LA PARTE DOVE SPIEGHI MAX, MIN, MON
+
+Assume that we want to state the following property: "\textit{Whenever there is a login by any user x, then eventually the user x logs out}". In LTL we can write the following formula: 
+\begin{align*}
+    & \square{(\text{(action = login)}\rightarrow \underline{\text{let}}\ \text{k} =\ userId\ \underline{\text{in}}\ \lozenge(\text{action = logout}\land userId = k))}
+\end{align*}
+In this formula we use the operator \underline{let} \_ \underline{in} \_ to bind the value of $userId$ in the current event to the local variable $k$ whenever $\text{(action = login)}$ in the current event; we then impose the condition that the value of $userId$ in some event in future must be same as the user id bound to k and that the action of the event must be logout.
+In EAGLE we can express the same property with the following rules:
+\begin{align*}
+    & \underline{\text{min}}\ \text{EvLogout}(\underline{\text{string}}\ k) = (\text{action = logout}\land userId = k) \lor \bigcirc \text{EvLogout}(k) \\
+    & \underline{\text{mon}}\ M_2 = ((\text{action = login})\rightarrow \text{EvLogout}(userId)) 
+\end{align*}
+As a result, rules in EAGLE give us the power to create specific temporal operators as well as to bind and modify data.
+This property turns out to be crucial for succinctly expressing executions of attack-safe systems.
+To make the paper self-contained, we go back to the syntax and semantics of EAGLE.
+\subsection{Syntax and Semantics}
+A specification $S$ is made up of an observer part $O$ and a declaration part $D$.
+$O$ is made up of zero or more monitor definitions $M$, which define what will be watched, while $D$ is made up of zero or more rule definitions $R$.
+The names of rules and monitors are $(N)$.
+\begin{align*}
+    S &::= D \, O \\
+    D &::= R^* \\
+    O &::= M^* \\
+    R &::= \{ \underline{\text{max}}\ |\ \underline{\text{min}} \}\ N(T_1\ x_1, \ldots, T_n\ x_n) = F \\
+    M &::= \underline{\text{mon}}\ N = F \\
+    T &::= \underline{\text{Form}}\ |\ \text{primitive type} \\
+    F &::= \text{expression}\ |\ \underline{\text{true}}\ |\ \underline{\text{false}}\ |\ \neg F\ |\ F_1 \land F_2\ |\ F_1 \lor F_2\ |\\
+            &F_1 \rightarrow F_2\ |\ \bigcirc F\ |\ \odot F\ |\ F_1 \cdot F_2\ |\ N(F_1, \ldots, F_n)\ |\ x_i
 \end{align*}

### 2023-08-31 Paper/main.tex
update

@@
 \begin{align*}
+& \underline{\text{min}}\ \text{Next}(\underline{\text{Form}}\ F) = \bigcirc F \\
 & \underline{\text{max}}\ \text{Always}(\underline{\text{Form}}\ F) = F \land \bigcirc \text{Always}(F) \\
-& \underline{\text{min}}\ \text{Eventually}(\underline{\text{Form}}\ F) = F \lor \bigcirc \text{Eventually}(F) 
+& \underline{\text{min}}\ \text{Eventually}(\underline{\text{Form}}\ F) = F \lor \bigcirc \text{Eventually}(F) \\
+& \underline{\text{min}}\ \text{Until}(\underline{\text{Form}}\ F_1, \underline{\text{Form}}\ F_2) = F_2 \lor (F_1 \land \bigcirc \text{Until}(F_1,F_2))
 \end{align*}
@@
     O &::= M^* \\
-    R &::= \{ \underline{\text{max}}\ |\ \underline{\text{min}} \}\ N(T_1\ x_1, \ldots, T_n\ x_n) = F \\
+    R &::= \{\underline{\text{max}}\ |\ \underline{\text{min}} \}\ N(T_1\ x_1, \ldots, T_n\ x_n) = F \\
     M &::= \underline{\text{mon}}\ N = F \\
@@
 \end{align*}
+A term indicating whether the interpretation is maximal ($\underline{\text{max}}$) or minimal ($\underline{\text{min}}$) comes before a rule specification $R$.
+While minimal rules establish \textit{liveness properties} (something good eventually happens), maximal rules define \textit{safety properties} (nothing bad ever happens); as we will see, the difference becomes important only when we are evaluating the formula in specific instants.
+Finally,the definitions starting with keyword $\underline{\text{mon}}$ specifies the EAGLE formulas to be monitored.
 

### 2023-08-31 Paper/main.tex
update

@@
 
+%% INTRODURRE VELOCE ESEMPIO E STRUTTURA DEL PROGRAMMA
+
 %%%%%   DA RISCRIVERE ALLA FINE
@@
 \end{align*}
+The types of parameters include primitive types like \underline{int}, \underline{long}, \underline{float}, etc. or formulas of type \underline{Form}.
 A term indicating whether the interpretation is maximal ($\underline{\text{max}}$) or minimal ($\underline{\text{min}}$) comes before a rule specification $R$.
 While minimal rules establish \textit{liveness properties} (something good eventually happens), maximal rules define \textit{safety properties} (nothing bad ever happens); as we will see, the difference becomes important only when we are evaluating the formula in specific instants.
-Finally,the definitions starting with keyword $\underline{\text{mon}}$ specifies the EAGLE formulas to be monitored.
+Recursive definitions of rules are permitted as long as they are tightly guarded by a temporal operator.
+Finally,the definitions starting with keyword $\underline{\text{mon}}$ specifies the EAGLE formulas to be monitored and cannot have a recursive definition.
+We shall see that monitored rules will evolve as new events appear.
 

### 2023-09-02 Paper/main.tex
update

@@
 Also, the $\text{Always}$ operator is defined as maximal solution of the equation $X=F\land \bigcirc X$ and the $\text{Sometime}$ operator represents the minimal solution to the equation $X=F\lor \bigcirc X$.
+
 %%  QUI MANCA LA PARTE DOVE SPIEGHI MAX, MIN, MON
@@
 \subsection{Syntax and Semantics}
+\textbf{Syntax.}
 A specification $S$ is made up of an observer part $O$ and a declaration part $D$.
@@
 
+\textbf{Semantics.} First of all two assumptions are important:\begin{enumerate}
+    \item There is a finite sequence of occurrences called $\sigma$ that is a merge of the system registered logs organized by ascending time. An example of event could be the following: 
+    \begin{align*}
+        & \text{LoginLogoutEvent}\{userId:\ \underline{\text{string}},action:\ \underline{\text{int}}, userId:\ \time{\text{double}}\} 
+    \end{align*}
+\end{enumerate}
+
+\subsection{Relationship to Other Logics}
+%% RAPPORTI CON LE ALTRE LOGICHE?????
+According to~\cite{barringer2004rule},
+
 We will use it to create formulas that describe the absence of some attacks and given a trace of events $\sigma$ generated from the logs we will check whether
+
+\section{MONID: Structure analysis}
+
+\section{Example of attacks}
+
+\section{Conclusions}
 

### 2023-09-02 Paper/main.tex
update

@@
 \textbf{Semantics.} First of all two assumptions are important:\begin{enumerate}
-    \item There is a finite sequence of occurrences called $\sigma$ that is a merge of the system registered logs organized by ascending time. An example of event could be the following: 
+    \item There is a finite sequence of occurrences called $\sigma=\alpha_{1},\ldots ,\alpha_{n}$ that is a merge of the system registered logs organized by ascending time. 
+    The structure of an event record $\alpha_i$ is the following:
     \begin{align*}
-        & \text{LoginLogoutEvent}\{userId:\ \underline{\text{string}},action:\ \underline{\text{int}}, userId:\ \time{\text{double}}\} 
+        & \text{LoginLogoutEvent}\{userId:\underline{\text{string}},action: \underline{\text{int}}, time: \underline{\text{double}}\} 
     \end{align*}
+    An example of event could be: $\{userId:\text{"Lorenzo"},action:\text{login},time:\text{20}\}$
 \end{enumerate}

### 2023-09-02 Paper/main.tex
update

@@
 To make the paper self-contained, we go back to the syntax and semantics of EAGLE.
+
+Lastly, two assumptions must be made:\begin{enumerate}
+    \item There is a finite sequence of occurrences called $\sigma=\alpha_{1},\ldots ,\alpha_{n}$ that is a merge of the system registered logs organized by ascending time. 
+    The structure of an event record $\alpha_i$ is the following:
+    \begin{align*}
+        & \text{LoginLogoutEvent}\{userId:\underline{\text{string}},action: \underline{\text{int}}, time: \underline{\text{double}}\} 
+    \end{align*}
+    An example of event could be: $\{userId:\text{"Lori"},action:\text{login},time:\text{20}\}$
+    \item For each attack, there is a formula $F$ which specifies the absence of it.
+\end{enumerate}
+
 \subsection{Syntax and Semantics}
@@
 
-\textbf{Semantics.} First of all two assumptions are important:\begin{enumerate}
-    \item There is a finite sequence of occurrences called $\sigma=\alpha_{1},\ldots ,\alpha_{n}$ that is a merge of the system registered logs organized by ascending time. 
-    The structure of an event record $\alpha_i$ is the following:
-    \begin{align*}
-        & \text{LoginLogoutEvent}\{userId:\underline{\text{string}},action: \underline{\text{int}}, time: \underline{\text{double}}\} 
-    \end{align*}
-    An example of event could be: $\{userId:\text{"Lorenzo"},action:\text{login},time:\text{20}\}$
-\end{enumerate}
+\textbf{Semantics.} The semantics of the logic is defined from the \textit{satisfaction} relation $\vDash$ which defines whether an execution trace $\sigma$ satisfy the specification $\varphi$.
+The $i$'th state $s_i$ of a trace $\sigma$ is denoted by $\sigma(i)$.
+The term $\sigma[i, j]$ denotes the sub-trace of $\sigma$ from position $i$ to position $j$, both positions included; if $i \geq j$ then $\sigma[i, j]$ denotes the empty trace.
+
+Given a trace $\sigma$ and a specification $D \, O$, satisfaction is defined as follows:
+\[
+\sigma \models D \, O \text{ iff } \forall (\underline{\text{mon}}\ N = F) \in O.\ \sigma ,1 \vDash_D \, F
+\]
+That is, a trace satisfies a specification if the trace, observed from position 1 (the first state), satisfies each monitored formula.
 

### 2023-09-02 Paper/main.tex
update

@@
 \section{EAGLE, a Temporal Monitoring Logic}
-According to~\cite{barringer2004program}, EAGLE offers a succinct but powerful set of primitives, essentially supporting recursive parameterized equations, with a minimal/maximal fix-point semantics together with three temporal operators: next-time ($\bigcirc$), previous-time ($\odot$), and concatenation ($\cdot$).
+According to~\cite{barringer2004program}, EAGLE offers a succinct but powerful set of primitives, essentially supporting recursive parameterized equations, with a minimal/maximal fix-point semantics together with three temporal operators: next-time ($\bigcirc$), previous-time ($\astrosun$), and concatenation ($\cdot$).
 \subsection{Basics}
@@
 
-\textbf{Semantics.} The semantics of the logic is defined from the \textit{satisfaction} relation $\vDash$ which defines whether an execution trace $\sigma$ satisfy the specification $\varphi$.
+\textbf{Semantics.} The semantics of the logic is defined from the \textit{satisfaction} relation $\models$ which defines whether an execution trace $\sigma$ satisfy the specification $\varphi=D \, O$.
 The $i$'th state $s_i$ of a trace $\sigma$ is denoted by $\sigma(i)$.
@@
 \[
-\sigma \models D \, O \text{ iff } \forall (\underline{\text{mon}}\ N = F) \in O.\ \sigma ,1 \vDash_D \, F
+\sigma \models D \, O \text{ iff } \forall (\underline{\text{mon}}\ N = F) \in O.\ \sigma ,1 \models_D \, F
 \]
-That is, a trace satisfies a specification if the trace, observed from position 1 (the first state), satisfies each monitored formula.
+In other words, a trace satisfies a specification if it fulfills each monitored formula when monitoring from position 1, which is the initial state.
+An \textbf{important} note is that, given a trace $\sigma=\alpha_{1},\ldots ,\alpha_{n}$ the index $i$ of the trace can become $0$ (before the first state) and $n+1$, thus going beyond the limits; both of these situations result in rule applications evaluating to either true if maximal or false if minimal, without first taking the rules body into account.
+The definition of the satisfaction relation $\models_D \subseteq (Trace \times \mathbb{N}) \times \underline{\text{Form}}$, for a set of rule definitions $D$, is the following:
+\begin{align*}
+    \sigma, i &\models_D\ expression &\text{iff} \quad &1 \leq i \leq |\sigma|\ \text{and}\ evaluate(expression)(\sigma(i)) = true \\
+    \sigma, i &\models_D\ \underline{\text{true}} \\
+    \sigma, i &\not\models_D\ \underline{\text{false}} \\
+    \sigma, i &\models_D \neg F &\text{iff} \quad &\sigma, i \not\models_D F \\
+    \sigma, i &\models_D F_1 \land F_2 &\text{iff} \quad &\sigma, i \models_D F_1\ \text{and}\ \sigma, i \models_D F_2 \\
+    \sigma, i &\models_D F_1 \lor F_2 &\text{iff} \quad &\sigma, i \models_D F_1\ \text{or}\ \sigma, i \models_D F_2 \\
+    \sigma, i &\models_D F_1 \rightarrow F_2 &\text{iff} \quad &\sigma, i \models_D F_1 \text{ implies } \sigma, i \models_D F_2 \\
+    \sigma, i &\models_D \bigcirc F &\text{iff} \quad &i \leq |\sigma|\ \text{and}\ \sigma, i+1 \models_D F \\
+    \sigma, i &\models_D \Bigg\bigodot F &\text{iff} \quad &1 \leq i\ \text{and}\ \sigma, i-1 \models_D F \\
+    \sigma, i &\models_D F_1 \cdot F_2 &\text{iff} \quad &\exists j \text{ such that } i \leq j \leq |\sigma|+1 \quad \text{and} \\
+    & &&\sigma[1, j-1], i \models_D F_1 \quad \text{and} \quad \sigma[j,|\sigma|], 1 \models_D F_2 \\
+    \sigma, i &\models_D N(F, P) &\text{iff} 
+    &\begin{cases}
+    \text{if } 1 \leq i \leq |\sigma| \text{ then:} \\
+    \sigma, i \models_D F[ f \mapsto F, p \mapsto \text{evaluate}(P)(\sigma(i))] \\
+    \text{where } (N(\text{Form } f, \text{T } p) = F) \in D \\
+    \text{otherwise, if } i = 0 \text{ or } i = |\sigma|+1 \text{ then:} \\
+    \text{rule } N \text{ is defined as max in } D
+    \end{cases}
+\end{align*}
 

### 2023-09-04 Paper/main.tex
update

@@
 As we can see, rules are parameterized by an EAGLE formula (of type $\underline{\text{Form}}$), which means that we will be able to write EAGLE formulas such as $\text{Always}(\text{Eventually}(x>0))$.
-Also, the $\text{Always}$ operator is defined as maximal solution of the equation $X=F\land \bigcirc X$ and the $\text{Sometime}$ operator represents the minimal solution to the equation $X=F\lor \bigcirc X$.
+Also, the $\text{Always}$ operator is defined as maximal solution of the equation $X=F\land \bigcirc X$ and the $\text{Eventually}$ operator represents the minimal solution to the equation $X=F\lor \bigcirc X$.
 
@@
     \sigma, i &\models_D \Bigg\bigodot F &\text{iff} \quad &1 \leq i\ \text{and}\ \sigma, i-1 \models_D F \\
-    \sigma, i &\models_D F_1 \cdot F_2 &\text{iff} \quad &\exists j \text{ such that } i \leq j \leq |\sigma|+1 \quad \text{and} \\
-    & &&\sigma[1, j-1], i \models_D F_1 \quad \text{and} \quad \sigma[j,|\sigma|], 1 \models_D F_2 \\
-    \sigma, i &\models_D N(F, P) &\text{iff} 
+    \sigma, i &\models_D F_1 \cdot F_2 &\text{iff} \quad &\exists j \text{ s.t.
+} i \leq j \leq |\sigma|+1\ \text{and}\ \sigma^{[1, j-1]}, i \models_D F_1\ \text{and}\ \sigma^{[j,|\sigma|]}, 1 \models_D F_2 \\
+    \sigma, i &\models_D N(\overline{F}, \overline{P}) &\text{iff} 
     &\begin{cases}
     \text{if } 1 \leq i \leq |\sigma| \text{ then:} \\
-    \sigma, i \models_D F[ f \mapsto F, p \mapsto \text{evaluate}(P)(\sigma(i))] \\
-    \text{where } (N(\text{Form } f, \text{T } p) = F) \in D \\
+    \quad \sigma, i \models_D F[ \overline{f} \mapsto \overline{F}, \overline{p} \mapsto\ evaluate(\overline{P})(\sigma(i))] \\
+    \quad \text{where } (N(\underline{\overline{\text{Form}}}\ \overline{f}, \text{T } p) = F) \in D \\
     \text{otherwise, if } i = 0 \text{ or } i = |\sigma|+1 \text{ then:} \\
-    \text{rule } N \text{ is defined as max in } D
+    \quad \text{rule } N \text{ is defined as }\underline{\text{max}}\text{ in } D
     \end{cases}

### 2023-09-04 Paper/main.tex
update

@@
 As we can see, rules are parameterized by an EAGLE formula (of type $\underline{\text{Form}}$), which means that we will be able to write EAGLE formulas such as $\text{Always}(\text{Eventually}(x>0))$.
-Also, the $\text{Always}$ operator is defined as maximal solution of the equation $X=F\land \bigcirc X$ and the $\text{Eventually}$ operator represents the minimal solution to the equation $X=F\lor \bigcirc X$.
+Also, the $\text{Always}$ operator is defined as maximal solution of the equation $X=F\land \bigcirc X$, while the $\text{Eventually}$ operator represents the minimal solution to the equation $X=F\lor \bigcirc X$.
 
@@
 \begin{align*}
-    & \square{(\text{(action = login)}\rightarrow \underline{\text{let}}\ \text{k} =\ userId\ \underline{\text{in}}\ \lozenge(\text{action = logout}\land userId = k))}
+    & \square{(\text{(action = login)}\rightarrow \underline{\text{let}}\ \text{k} =userId\ \underline{\text{in}}\ \lozenge(\text{action = logout}\land userId = k))}
 \end{align*}

### 2023-09-04 Paper/main.tex
update

@@
 
-%%  QUI MANCA LA PARTE DOVE SPIEGHI MAX, MIN, MON
-
 Assume that we want to state the following property: "\textit{Whenever there is a login by any user x, then eventually the user x logs out}". In LTL we can write the following formula: 
@@
 This property turns out to be crucial for succinctly expressing executions of attack-safe systems.
-To make the paper self-contained, we go back to the syntax and semantics of EAGLE.
+As we can see, each rule begins with a term that describes its type ($\underline{\text{min}}$, $\underline{\text{max}}$, $\underline{\text{mon}}$); we will read more about it later.
 
 Lastly, two assumptions must be made:\begin{enumerate}
-    \item There is a finite sequence of occurrences called $\sigma=\alpha_{1},\ldots ,\alpha_{n}$ that is a merge of the system registered logs organized by ascending time. 
+    \item There is a finite sequence of events called $\sigma=\alpha_{1},\ldots ,\alpha_{n}$ that is a merge of the system registered logs organized by ascending time. 
     The structure of an event record $\alpha_i$ is the following:
@@
 \end{enumerate}
+To make the paper self-contained, we will present the syntax and semantics of EAGLE.
 
@@
 \end{align*}
-The types of parameters include primitive types like \underline{int}, \underline{long}, \underline{float}, etc. or formulas of type \underline{Form}.
+The types $T_i$ of parameters include primitive types like \underline{int}, \underline{long}, \underline{float}, etc.
+or formulas of type \underline{Form}.
+Definitions starting with keyword $\underline{\text{mon}}$ specifies the EAGLE formulas to be monitored and cannot have a recursive definition.
+We will see that these kind of rules will evolve as new events appear.
 A term indicating whether the interpretation is maximal ($\underline{\text{max}}$) or minimal ($\underline{\text{min}}$) comes before a rule specification $R$.
-While minimal rules establish \textit{liveness properties} (something good eventually happens), maximal rules define \textit{safety properties} (nothing bad ever happens); as we will see, the difference becomes important only when we are evaluating the formula in specific instants.
-Recursive definitions of rules are permitted as long as they are tightly guarded by a temporal operator.
-Finally,the definitions starting with keyword $\underline{\text{mon}}$ specifies the EAGLE formulas to be monitored and cannot have a recursive definition.
-We shall see that monitored rules will evolve as new events appear.
+While minimal rules establish \textit{liveness properties} (something good eventually happens), maximal rules define \textit{safety properties} (nothing bad ever happens); we shall see that the difference becomes important only when we are evaluating the at the boundaries of a trace.
+Finally, recursive definitions of rules are permitted as long as they are tightly guarded by a temporal operator.
 
-\textbf{Semantics.} The semantics of the logic is defined from the \textit{satisfaction} relation $\models$ which defines whether an execution trace $\sigma$ satisfy the specification $\varphi=D \, O$.
+\textbf{Semantics.} The semantics of the logic is defined in terms of the \textit{satisfaction} relation $\models$ which defines whether an execution trace $\sigma$ satisfy the specification $\varphi=D \, O$.
 The $i$'th state $s_i$ of a trace $\sigma$ is denoted by $\sigma(i)$.
@@
 \]
-In other words, a trace satisfies a specification if it fulfills each monitored formula when monitoring from position 1, which is the initial state.
+In other words, a trace satisfies a specification if it fulfills each monitored formula ($\underline{\text{mon}}$) when monitoring from position 1, which is the initial state.
 An \textbf{important} note is that, given a trace $\sigma=\alpha_{1},\ldots ,\alpha_{n}$ the index $i$ of the trace can become $0$ (before the first state) and $n+1$, thus going beyond the limits; both of these situations result in rule applications evaluating to either true if maximal or false if minimal, without first taking the rules body into account.
+Let us explain why using an example.
+Given the definition
+\begin{align*}
+    & \underline{\text{min}}\ \text{Property}(\underline{\text{Form}}\ F) =F \lor \bigcirc \text{Property}(F)
+\end{align*}
+and the sequence $\sigma=\alpha_{1},\ldots ,\alpha_{n}$, Property will evaluate to \textit{true} 
+\begin{align*}
+    & evaluate(\ldots evaluate(evaluate(\text{Property},\alpha_1),\alpha_2),\ldots,\alpha_n)=true
+\end{align*}
+if and only if the rule is true at some given event $\alpha_i$. 
+Once the sequence has been completely analyzed, we will obtain a big disjunction (in case of a \underline{\text{min}}) or a big conjunction (in case of a \underline{\text{max}}) and the final value will close the evalutation.
+
 The definition of the satisfaction relation $\models_D \subseteq (Trace \times \mathbb{N}) \times \underline{\text{Form}}$, for a set of rule definitions $D$, is the following:
 \begin{align*}
-    \sigma, i &\models_D\ expression &\text{iff} \quad &1 \leq i \leq |\sigma|\ \text{and}\ evaluate(expression)(\sigma(i)) = true \\
+    \sigma, i &\models_D\ expression &\text{iff} \quad &1 \leq i \leq |\sigma|\ \text{and}\ evaluate(expression, \sigma(i)) = true \\
     \sigma, i &\models_D\ \underline{\text{true}} \\
@@
     \text{if } 1 \leq i \leq |\sigma| \text{ then:} \\
-    \quad \sigma, i \models_D F[ \overline{f} \mapsto \overline{F}, \overline{p} \mapsto\ evaluate(\overline{P})(\sigma(i))] \\
+    \quad \sigma, i \models_D F[ \overline{f} \mapsto \overline{F}, \overline{p} \mapsto\ evaluate(\overline{P}, \sigma(i))] \\
     \quad \text{where } (N(\underline{\overline{\text{Form}}}\ \overline{f}, \text{T } p) = F) \in D \\

### 2023-09-04 Paper/main.tex
update

@@
 
-\textbf{Semantics.} The semantics of the logic is defined in terms of the \textit{satisfaction} relation $\models$ which defines whether an execution trace $\sigma$ satisfy the specification $\varphi=D \, O$.
+\textbf{Semantics.} The semantics of the logic is defined in terms of the \textit{satisfaction} relation $\models$ which defines whether a finite execution trace $\sigma$ satisfy the specification $\varphi=D \, O$.
 The $i$'th state $s_i$ of a trace $\sigma$ is denoted by $\sigma(i)$.
@@
 In other words, a trace satisfies a specification if it fulfills each monitored formula ($\underline{\text{mon}}$) when monitoring from position 1, which is the initial state.
+The definition of the satisfaction relation $\models_D \subseteq (Trace \times \mathbb{N}) \times \underline{\text{Form}}$, for a set of rule definitions $D$, is the following:
+\begin{align*}
+    \sigma, i &\models_D\ expression &\text{iff} \quad &1 \leq i \leq |\sigma|\ \text{and}\ evaluate(expression, \sigma(i)) = true \\
+    \sigma, i &\models_D\ \underline{\text{true}} \\
+    \sigma, i &\not\models_D\ \underline{\text{false}} \\
+    \sigma, i &\models_D \neg F &\text{iff} \quad &\sigma, i \not\models_D F \\
+    \sigma, i &\models_D F_1 \land F_2 &\text{iff} \quad &\sigma, i \models_D F_1\ \text{and}\ \sigma, i \models_D F_2 \\
+    \sigma, i &\models_D F_1 \lor F_2 &\text{iff} \quad &\sigma, i \models_D F_1\ \text{or}\ \sigma, i \models_D F_2 \\
+    \sigma, i &\models_D F_1 \rightarrow F_2 &\text{iff} \quad &\sigma, i \models_D F_1 \text{ implies } \sigma, i \models_D F_2 \\
+    \sigma, i &\models_D \bigcirc F &\text{iff} \quad &i \leq |\sigma|\ \text{and}\ \sigma, i+1 \models_D F \\
+    \sigma, i &\models_D \odot F &\text{iff} \quad &1 \leq i\ \text{and}\ \sigma, i-1 \models_D F \\
+    \sigma, i &\models_D F_1 \cdot F_2 &\text{iff} \quad &\exists j \text{ s.t.
+} i \leq j \leq |\sigma|+1\ \text{and}\ \sigma^{[1, j-1]}, i \models_D F_1\ \text{and}\ \sigma^{[j,|\sigma|]}, 1 \models_D F_2 \\
+    \sigma, i &\models_D N(\overline{F}, \overline{P}) &\text{iff} 
+    &\begin{cases}
+    \text{if } 1 \leq i \leq |\sigma| \text{ then:} \\
+    \quad \sigma, i \models_D F[ \overline{f} \mapsto \overline{F}, \overline{p} \mapsto\ evaluate(\overline{P}, \sigma(i))] \\
+    \quad \text{where } (N(\underline{\overline{\text{Form}}}\ \overline{f}, \text{T } p) = F) \in D \\
+    \text{otherwise, if } i = 0 \text{ or } i = |\sigma|+1 \text{ then:} \\
+    \quad \text{rule } N \text{ is defined as }\underline{\text{max}}\text{ in } D
+    \end{cases}
+\end{align*}
+where $\overline{F}$ and $\overline{P}$ represent tuples of type $\overline{\text{Form}}$ and $\overline{T}$ respectively
+
 An \textbf{important} note is that, given a trace $\sigma=\alpha_{1},\ldots ,\alpha_{n}$ the index $i$ of the trace can become $0$ (before the first state) and $n+1$, thus going beyond the limits; both of these situations result in rule applications evaluating to either true if maximal or false if minimal, without first taking the rules body into account.
@@
 Once the sequence has been completely analyzed, we will obtain a big disjunction (in case of a \underline{\text{min}}) or a big conjunction (in case of a \underline{\text{max}}) and the final value will close the evalutation.
-
-The definition of the satisfaction relation $\models_D \subseteq (Trace \times \mathbb{N}) \times \underline{\text{Form}}$, for a set of rule definitions $D$, is the following:
+Continuing the example, the extended formula will be:
 \begin{align*}
-    \sigma, i &\models_D\ expression &\text{iff} \quad &1 \leq i \leq |\sigma|\ \text{and}\ evaluate(expression, \sigma(i)) = true \\
-    \sigma, i &\models_D\ \underline{\text{true}} \\
-    \sigma, i &\not\models_D\ \underline{\text{false}} \\
-    \sigma, i &\models_D \neg F &\text{iff} \quad &\sigma, i \not\models_D F \\
-    \sigma, i &\models_D F_1 \land F_2 &\text{iff} \quad &\sigma, i \models_D F_1\ \text{and}\ \sigma, i \models_D F_2 \\
-    \sigma, i &\models_D F_1 \lor F_2 &\text{iff} \quad &\sigma, i \models_D F_1\ \text{or}\ \sigma, i \models_D F_2 \\
-    \sigma, i &\models_D F_1 \rightarrow F_2 &\text{iff} \quad &\sigma, i \models_D F_1 \text{ implies } \sigma, i \models_D F_2 \\
-    \sigma, i &\models_D \bigcirc F &\text{iff} \quad &i \leq |\sigma|\ \text{and}\ \sigma, i+1 \models_D F \\
-    \sigma, i &\models_D \Bigg\bigodot F &\text{iff} \quad &1 \leq i\ \text{and}\ \sigma, i-1 \models_D F \\
-    \sigma, i &\models_D F_1 \cdot F_2 &\text{iff} \quad &\exists j \text{ s.t.
-} i \leq j \leq |\sigma|+1\ \text{and}\ \sigma^{[1, j-1]}, i \models_D F_1\ \text{and}\ \sigma^{[j,|\sigma|]}, 1 \models_D F_2 \\
-    \sigma, i &\models_D N(\overline{F}, \overline{P}) &\text{iff} 
-    &\begin{cases}
-    \text{if } 1 \leq i \leq |\sigma| \text{ then:} \\
-    \quad \sigma, i \models_D F[ \overline{f} \mapsto \overline{F}, \overline{p} \mapsto\ evaluate(\overline{P}, \sigma(i))] \\
-    \quad \text{where } (N(\underline{\overline{\text{Form}}}\ \overline{f}, \text{T } p) = F) \in D \\
-    \text{otherwise, if } i = 0 \text{ or } i = |\sigma|+1 \text{ then:} \\
-    \quad \text{rule } N \text{ is defined as }\underline{\text{max}}\text{ in } D
-    \end{cases}
+    & F \lor \bigcirc(F \lor \bigcirc(\ldots F \lor \bigcirc \textbf{Property(F)})) 
 \end{align*}
+and the bold \textbf{Property(F)} will be evaluated to false, since it is the last of a big disjunction.
+\subsection{Evaluation algorithm}
+As already mentioned, monitored formulas ($\underline{\text{mon}}$) evolve at every event, to store relevant informations about past istants.
+The evolved formula's value is determined at the conclusion of the event series; if it is true, the formula is satisfied by the event sequence; otherwise, it is violated.
+Formally, a formula $F' = \text evaluate(F, \alpha_i)$ is created when a formula $F$ is evaluated at an event $\alpha_i = \sigma(i)$ and has the condition that $\sigma, i \models F$ if and only if $\sigma, i+1 \models F'$.
+With $F$ being the evolved formula, we compute the boolean function $value(F)$ at the conclusion of the trace.
+The definition of the function $evaluate: \underline{\text{Form}} \times \underline{\text{State}} \to \underline{\text{Form}}$ uses an auxiliary function $update: \underline{\text{Form}} \times \underline{\text{State}} \to \underline{\text{Form}}$ to pre-evaluate a formula if it is guarded by the previous operator $\bigodot$.
 

### 2023-09-04 Paper/main.tex
update

@@
 
+The definitions of $evaluate$, $update$ and $value$ on the different primitive operators are the following:
+\begin{align*}
+    evaluate(\underline{\text{true}},\alpha_i) & =\ \underline{\text{true}} \\
+    evaluate(\underline{\text{false}},\alpha_i) & =\ \underline{\text{false}} \\ 
+    evaluate(\text{exp},\alpha_i) & =\ \text{value of } exp\text{ in } \alpha_i \\
+    evaluate(F_1\ op\ F_2,\alpha_i) & =\  evaluate(F_1,\alpha_i)\ op\ evaluate(F_2,\alpha_i)\quad \text{where op can be }\lor , \land, \rightarrow \\
+    evaluate(\neg{F},\alpha_i) & =\ \neg{evaluate(F,\alpha_i)}  \\
+    evaluate(\bigcirc F,\alpha_i) & =\ update(F,\alpha_i)  \\
+    evaluate(F_1 \cdot F_2,\alpha_i) & =
+    \begin{cases}
+        \text{if } value(F_1)=\ \underline{\text{false}} \\
+        \quad  evaluate(F_1,\alpha_i)\cdot F_2 \\
+        \text{else} \\
+        \quad  evaluate(F_1,\alpha_i)\cdot F_2 \lor evaluate(F_2,\alpha_i)
+    \end{cases}
+\end{align*}
+
 \subsection{Relationship to Other Logics}

### 2023-09-04 Paper/main.tex
update

@@
 
-The definitions of $evaluate$, $update$ and $value$ on the different primitive operators are the following:
+According to~\cite{barringer2004rule}, the definitions of $evaluate$, $update$ and $value$ on the different primitive operators are the following ($op\text{ can be }\lor , \land, \rightarrow $):
+%evaluate
+\begin{align*}
+    evaluate(\underline{\text{true}},\alpha_i) & =\ \underline{\text{true}} \\
+    evaluate(\underline{\text{false}},\alpha_i) & =\ \underline{\text{false}} \\ 
+    evaluate(\text{exp},\alpha_i) & =\ \text{value of } exp\text{ in } \alpha_i \\
+    evaluate(F_1\ op\ F_2,\alpha_i) & =\  evaluate(F_1,\alpha_i)\ op\ evaluate(F_2,\alpha_i)\\
+    evaluate(\neg{F},\alpha_i) & =\ \neg{evaluate(F,\alpha_i)}  \\
+    evaluate(\bigcirc F,\alpha_i) & =\ update(F,\alpha_i)  \\
+    evaluate(F_1 \cdot F_2,\alpha_i) & =
+    \begin{cases}
+        evaluate(F_1,\alpha_i)\cdot F_2 \quad \text{if } value(F_1)=\ \underline{\text{false}} \\
+        evaluate(F_1,\alpha_i)\cdot F_2 \lor evaluate(F_2,\alpha_i)\quad \text{else} 
+    \end{cases}\\ \\
+%value
+    value(\underline{\text{true}}) & =\ \underline{\text{true}} \\
+    value(\underline{\text{false}}) & =\ \underline{\text{false}} \\ 
+    value(\text{exp}) & =\ \underline{\text{false}} \\
+    value(F_1\ op\ F_2) & =\  value(F_1)\ op\ value(F_2) \\
+    value(\neg{F}) & =\ \neg{value(F)}  \\
+    value(\bigcirc F) & =
+    \begin{cases}
+        F \quad \text{if at the beginning of a trace} \\
+        \underline{\text{false}}\quad \text{if at the end of a trace} \\
+    \end{cases}\\
+    value(F_1 \cdot F_2,\alpha_i) & =\  value(F_1)\land value(F_2) 
+\end{align*}
+%update
 \begin{align*}

### 2023-09-04 Paper/main.tex
update

--- 2023-09-04 Paper/main.tex
@@
 
+According to~\cite{barringer2004rule}, the definitions of $evaluate$, $update$ and $value$ on the different primitive operators are the following ($op\text{ can be }\lor , \land, \rightarrow $):
+%evaluate
+\begin{align*}
+    evaluate(\underline{\text{true}},\alpha_i) & =\ \underline{\text{true}} \\
+    evaluate(\underline{\text{false}},\alpha_i) & =\ \underline{\text{false}} \\ 
+    evaluate(exp,\alpha_i) & =\ \text{value of } exp\text{ in } \alpha_i \\
+    evaluate(F_1\ op\ F_2,\alpha_i) & =\  evaluate(F_1,\alpha_i)\ op\ evaluate(F_2,\alpha_i)\\
+    evaluate(\neg{F},\alpha_i) & =\ \neg{evaluate(F,\alpha_i)}  \\
+    evaluate(\bigcirc F,\alpha_i) & =\ update(F,\alpha_i)  \\
+    evaluate(F_1 \cdot F_2,\alpha_i) & =
+    \begin{cases}
+        evaluate(F_1,\alpha_i)\cdot F_2 \quad \text{if } value(F_1)=\ \underline{\text{false}} \\
+        evaluate(F_1,\alpha_i)\cdot F_2 \lor evaluate(F_2,\alpha_i)\quad \text{else} 
+    \end{cases}\\ \\
+%value
+    value(\underline{\text{true}}) & =\ \underline{\text{true}} \\
+    value(\underline{\text{false}}) & =\ \underline{\text{false}} \\ 
+    value(exp) & =\ \underline{\text{false}} \\
+    value(F_1\ op\ F_2) & =\  value(F_1)\ op\ value(F_2) \\
+    value(\neg{F}) & =\ \neg{value(F)}  \\
+    value(\bigcirc F) & =
+    \begin{cases}
+        F \quad \text{if at the beginning of a trace} \\
+        \underline{\text{false}}\quad \text{if at the end of a trace} \\
+    \end{cases}\\
+    value(F_1 \cdot F_2,\alpha_i) & =\  value(F_1)\land value(F_2) \\ \\ 
+%update
+    update(\underline{\text{true}},\alpha_i) & =\ \underline{\text{true}} \\
+    update(\underline{\text{false}},\alpha_i) & =\ \underline{\text{false}} \\ 
+    update(exp,\alpha_i) & =\ exp\\
+    update(F_1\ op\ F_2,\alpha_i) & =\  update(F_1,\alpha_i)\ op\ update(F_2,\alpha_i) \\
+    update(\neg{F},\alpha_i) & =\ \neg{update(F,\alpha_i)}  \\
+    update(\bigcirc F,\alpha_i) & =\ update(F,\alpha_i)  \\
+    update(F_1 \cdot F_2,\alpha_i) & =\ update(F_1,\alpha_i)\cdot F_2
+\end{align*}
+
 \subsection{Relationship to Other Logics}

### 2023-09-04 Paper/main.tex
update

@@
         evaluate(F_1,\alpha_i)\cdot F_2 \lor evaluate(F_2,\alpha_i)\quad \text{else} 
-    \end{cases}\\ \\
-%value
-    value(\underline{\text{true}}) & =\ \underline{\text{true}} \\
-    value(\underline{\text{false}}) & =\ \underline{\text{false}} \\ 
-    value(exp) & =\ \underline{\text{false}} \\
-    value(F_1\ op\ F_2) & =\  value(F_1)\ op\ value(F_2) \\
-    value(\neg{F}) & =\ \neg{value(F)}  \\
-    value(\bigcirc F) & =
+    \end{cases}
+\end{align*}
+
+\noindent\begin{minipage}{0.3\linewidth}
+    \begin{flalign*}
+        \text{value}(\underline{\text{true}}) & = \underline{\text{true}} \\
+        \text{value}(\underline{\text{false}}) & = \underline{\text{false}} \\ 
+        \text{value}(\text{exp}) & = \underline{\text{false}} \\
+        \text{value}(F_1\ \text{op}\ F_2) & = \text{value}(F_1)\ \text{op}\ \text{value}(F_2) \\
+        \text{value}(\neg{F}) & = \neg{\text{value}(F)}  \\
+        \text{value}(\bigcirc F) & =
         \begin{cases}
-        F \quad \text{if at the beginning of a trace} \\
-        \underline{\text{false}}\quad \text{if at the end of a trace} \\
+            F & \text{if at the beginning of a trace} \\
+            \underline{\text{false}} & \text{if at the end of a trace} \\
         \end{cases}\\
-    value(F_1 \cdot F_2,\alpha_i) & =\  value(F_1)\land value(F_2) \\ \\ 
-%update
-    update(\underline{\text{true}},\alpha_i) & =\ \underline{\text{true}} \\
-    update(\underline{\text{false}},\alpha_i) & =\ \underline{\text{false}} \\ 
-    update(exp,\alpha_i) & =\ exp\\
-    update(F_1\ op\ F_2,\alpha_i) & =\  update(F_1,\alpha_i)\ op\ update(F_2,\alpha_i) \\
-    update(\neg{F},\alpha_i) & =\ \neg{update(F,\alpha_i)}  \\
-    update(\bigcirc F,\alpha_i) & =\ update(F,\alpha_i)  \\
-    update(F_1 \cdot F_2,\alpha_i) & =\ update(F_1,\alpha_i)\cdot F_2
+        \text{value}(F_1 \cdot F_2,\alpha_i) & = \text{value}(F_1)\land \text{value}(F_2)
+    \end{flalign*}
+\end{minipage}
+\hfill
+\begin{minipage}{0.5\linewidth}
+    \begin{align*}
+        \text{update}(\underline{\text{true}},\alpha_i) & = \underline{\text{true}} \\
+        \text{update}(\underline{\text{false}},\alpha_i) & = \underline{\text{false}} \\ 
+        \text{update}(\text{exp},\alpha_i) & = \text{exp}\\
+        \text{update}(F_1\ \text{op}\ F_2,\alpha_i) & = \text{update}(F_1,\alpha_i)\ \text{op}\ \text{update}(F_2,\alpha_i) \\
+        \text{update}(\neg{F},\alpha_i) & = \neg{\text{update}(F,\alpha_i)}  \\
+        \text{update}(\bigcirc F,\alpha_i) & = \text{update}(F,\alpha_i)  \\
+        \text{update}(F_1 \cdot F_2,\alpha_i) & = \text{update}(F_1,\alpha_i)\cdot F_2
     \end{align*}
+\end{minipage}
 

### 2023-09-04 Paper/main.tex
update

@@
         evaluate(F_1,\alpha_i)\cdot F_2 \lor evaluate(F_2,\alpha_i)\quad \text{else} 
-    \end{cases}
-\end{align*}
-
-\noindent\begin{minipage}{0.3\linewidth}
-    \begin{flalign*}
+    \end{cases} \\ \\
+%\end{align*}
+%\noindent\begin{minipage}[l]{\textwidth}
+   % \raggedright
+    %\begin{align*}
         \text{value}(\underline{\text{true}}) & = \underline{\text{true}} \\
@@
         \end{cases}\\
-        \text{value}(F_1 \cdot F_2,\alpha_i) & = \text{value}(F_1)\land \text{value}(F_2)
-    \end{flalign*}
-\end{minipage}
-\hfill
-\begin{minipage}{0.5\linewidth}
-    \begin{align*}
+        \text{value}(F_1 \cdot F_2,\alpha_i) & = \text{value}(F_1)\land \text{value}(F_2) \\ \\
+   % \end{align*}
+%\end{minipage}
+%\hfill
+%\begin{minipage}[r]{\textwidth}
+   % \raggedleft
+    %\begin{align*}
         \text{update}(\underline{\text{true}},\alpha_i) & = \underline{\text{true}} \\
@@
     \end{align*}
-\end{minipage}
+%\end{minipage}
+Let us introduce an example to better understand the algorithm. Given the following rules 
+\begin{align*}
+    &\underline{\text{max}}\ \text{Always}(\underline{\text{Form}}\ F)  = F \land \bigcirc \text{Always}(F) \\
+    &\underline{\text{min}}\ \text{EvTimedLogout}(\underline{\text{string }} k, \underline{\text{double }} t, \underline{\text{double }} \delta) = (\text{time} - t \leq \delta) \\
+    &\quad\quad\quad\quad \land ((action = \text{logout} \land userId = k) \lor \bigcirc \text{EvTimedLogout}(k, t, \delta)) \\
+    &\underline{\text{mon}}\ M_3  = \text{Always}((action = \text{login}) \rightarrow \text{EvTimedLogout}(userId,time, 100))
+\end{align*}
+and a trace $\sigma=\alpha_1,\alpha_2$, where $\alpha_1=\{userId:\text{"Lori"},action:\text{login},time:\text{17.0}\}$ and $\alpha_1=\{userId:\text{"Lori"},action:\text{logout},time:\text{150.0}\}$, the initial formula $M_3$ gets modified.
+At time $1$, when $\alpha_1$ is analyzed, $M_3$ changes as follows:
+\begin{align*}
+    &evaluate(M_3,\alpha_1)  = \text{EvTimedLogout("Lori",17.0, 100)} \land \\ 
+    &\quad\quad\quad\quad \text{Always}((action = \text{login}) \rightarrow \text{EvTimedLogout}(userId,time, 100))
+\end{align*}
+At the second event the predicate $(time-t \leq \delta)$ gets instantiated to $(150.0-17.0 \leq 100)$ which is false; thus the whole formula becomes false, indicating that the two-event trace violates the property $M_3$.
+
+It's essential to emphasize that the logic EAGLE operates with a \textit{finite} trace semantics; the monitoring algorithm, as written before, can check whetet the formula is satisfied or not only at the conclusion of a trace.
+However, in intrusion detection scenarios, the notion of the 'end of a trace' makes no sense since event sequences can theoretically be infinite.
+In such cases, the aim is to trigger an alarm as soon as a property is violated and this is done by continuously checking the formula's satisfaction status after each event.
 

### 2023-09-05 Paper/main.tex
update

@@
 
-According to~\cite{barringer2004rule}, the definitions of $evaluate$, $update$ and $value$ on the different primitive operators are the following ($op\text{ can be }\lor , \land, \rightarrow $):
+\textbf{Calculus.} According to~\cite{barringer2004rule}, the definitions of $evaluate$, $update$ and $value$ on the different primitive operators are the following ($op\text{ can be }\lor , \land, \rightarrow $):
 %evaluate
@@
 In such cases, the aim is to trigger an alarm as soon as a property is violated and this is done by continuously checking the formula's satisfaction status after each event.
+In practice, once we have the specification of the attack pattern in terms of $\varphi$, we can express the system's safe behavior as $\square(\neg \varphi)$.
+Note that if a sequence of events corresponding to the attack is discovered, this formula becomes false.
 

### 2023-09-05 Paper/main.tex
update

--- 2023-09-04 Paper/main.tex
@@
 The definition of the function $evaluate: \underline{\text{Form}} \times \underline{\text{State}} \to \underline{\text{Form}}$ uses an auxiliary function $update: \underline{\text{Form}} \times \underline{\text{State}} \to \underline{\text{Form}}$ to pre-evaluate a formula if it is guarded by the previous operator $\bigodot$.
+At the end (or at the beginning) of a trace, the
+function $value : \underline{\text{Form}} \rightarrow \{ \underline{\text{true}}, \underline{\text{false}} \}$, when applied to $F$, returns true if and only if $\sigma,|\sigma|+1 \models F$ (or $\sigma,0 \models F$), and returns false otherwise.
 
-According to~\cite{barringer2004rule}, the definitions of $evaluate$, $update$ and $value$ on the different primitive operators are the following ($op\text{ can be }\lor , \land, \rightarrow $):
+\textbf{Calculus.} According to~\cite{barringer2004rule}, the $evaluate$, $update$ and $value$ functions are defined a priori for all operators except for the
+rule application, for which get generated based on the definition of rules in the specification.
+The definitions of these functions on the different primitive operators are the following ($op\text{ can be }\lor , \land, \rightarrow $):
 %evaluate
+{\allowdisplaybreaks
 \begin{align*}
@@
     \end{align*}
-\end{minipage}
+    }
+%\end{minipage}
+Let us introduce an example to better understand the algorithm. Given the following rules 
+\begin{align*}
+    &\underline{\text{max}}\ \text{Always}(\underline{\text{Form}}\ F)  = F \land \bigcirc \text{Always}(F) \\
+    &\underline{\text{min}}\ \text{EvTimedLogout}(\underline{\text{string }} k, \underline{\text{double }} t, \underline{\text{double }} \delta) = (\text{time} - t \leq \delta) \\
+    &\quad\quad\quad\quad \land ((action = \text{logout} \land userId = k) \lor \bigcirc \text{EvTimedLogout}(k, t, \delta)) \\
+    &\underline{\text{mon}}\ M_3  = \text{Always}((action = \text{login}) \rightarrow \text{EvTimedLogout}(userId,time, 100))
+\end{align*}
+and a trace $\sigma=\alpha_1,\alpha_2$, where $\alpha_1=\{userId:\text{"Lori"},action:\text{login},time:\text{17.0}\}$ and $\alpha_1=\{userId:\text{"Lori"},action:\text{logout},time:\text{150.0}\}$, the initial formula $M_3$ gets modified.
+At time $1$, when $\alpha_1$ is analyzed, $M_3$ changes as follows:
+\begin{align*}
+    &evaluate(M_3,\alpha_1)  = \text{EvTimedLogout("Lori",17.0, 100)} \\ 
+    &\quad\quad\quad\quad \land\text{Always}((action = \text{login}) \rightarrow \text{EvTimedLogout}(userId,time, 100))
+\end{align*}
+At the second event the predicate $(time-t \leq \delta)$ gets instantiated to $(150.0-17.0 \leq 100)$ which is false; thus the whole formula becomes false, indicating that the two-event trace violates the property $M_3$.
+
+It's essential to emphasize that the logic EAGLE operates with a \textit{finite} trace semantics; the monitoring algorithm, as written before, can check whetet the formula is satisfied or not only at the conclusion of a trace.
+However, in intrusion detection scenarios, the notion of the "end of a trace" makes no sense since event sequences can theoretically be infinite.
+In such cases, the aim is to trigger an alarm as soon as a property is violated and this is done by continuously checking the formula's satisfaction status after each event.
+In practice, once we have the specification of the attack pattern in terms of $\varphi$, we can express the system's safe behavior as $\square(\neg \varphi)$.
+Note that if a sequence of events corresponding to the attack is discovered, this formula becomes false.
 

### 2023-09-05 Paper/main.tex
update

@@
 According to~\cite{barringer2004program}, EAGLE offers a succinct but powerful set of primitives, essentially supporting recursive parameterized equations, with a minimal/maximal fix-point semantics together with three temporal operators: next-time ($\bigcirc$), previous-time ($\astrosun$), and concatenation ($\cdot$).
-\subsection{Basics}
-In EAGLE recursion definitions are supported; in the current framework we can build the following definitions:
+\subsection{Basics}\label{sec:ltl}
+In EAGLE recursion definitions are supported; as exmaple, in the current framework we can build the following definitions:
 \begin{align*}
@@
 
-\subsection{Relationship to Other Logics}
-%% RAPPORTI CON LE ALTRE LOGICHE?????
-According to~\cite{barringer2004rule},
-
-We will use it to create formulas that describe the absence of some attacks and given a trace of events $\sigma$ generated from the logs we will check whether
-
+\subsection{Relationship to LTL+P}
+According to~\cite{barringer2004rule,barringer2003eagle}, we can define both \textbf{Future Time LTL} and \textbf{Past Time LTL} operators.
+The future modalities are written in section~\ref{sec:ltl}, therefore let us list the past ones:
+\begin{align*}
+    & \underline{\text{min}}\ \text{Previous}(\underline{\text{Form}}\ F) = \odot F \\
+    & \underline{\text{max}}\ \text{AlwaysInPast}(\underline{\text{Form}}\ F) = F \land \odot \text{AlwaysInPast}(F) \\
+    & \underline{\text{min}}\ \text{EventuallyInPast}(\underline{\text{Form}}\ F) = F \lor \odot \text{EventuallyInPast}(F) \\
+    & \underline{\text{min}}\ \text{Since}(\underline{\text{Form}}\ F_1, \underline{\text{Form}}\ F_2) = F_2 \lor (F_1 \land \odot \text{Since}(F_1,F_2))
+\end{align*}
+% Note that UNLESS is maximal
+By combining the definitions for the future and past time LTLs defined above, we obtain a temporal logic over the future, present and past, in which one can freely intermix the future and past time modalities.
 \section{MONID: Structure analysis}

### 2023-09-05 Paper/main.tex
update

@@
 \section{MONID: Structure analysis}
+MONID is a prototype which can detect intrusions either online or offline.
+System-level event data is delivered to MONID server either directly from relevant log files (offline) or online via application code that has been properly instrumented; to create a single event trace, the server combines events from numerous sources according to timestamps and preprocesses them into an abstract intermediate form.
+Note that sometimes it is necessary to watch events from multiple sources in order to detect certain attacks; then, MONID monitor checks this event trail against a predetermined criteria, and if the specification is broken, an intrusion alarm is raised.
+The entire framework is represented in figure~.
 

### 2023-09-05 Paper/main.tex
update

@@
 The logic allows us to describe temporal patterns involving reasoning about data-values observed in individual events.
-Unlike LTL, EAGLE allowa us to design attacks with fundamentally statistical characteristics; password guessing attacks and ICMP-flood denial of service attacks are two examples.
+Unlike LTL, EAGLE allows us to design attacks with fundamentally statistical characteristics; password guessing attacks and ICMP-flood denial of service attacks are two examples.
+We will monitor EAGLE formulas using an online algorithm that processes each event as it happens and updates the monitored formula to store a relevant summary.
+An intrusion alarm is triggered if, for any reason, the updated formula turns out to be false; as a result, the entire process operates in real-time.
 We'll see how the implementation, MONID, examines the event stream to see if the monitored formula (which is used to hold the pertinent summary of the system) is being broken.
@@
 and the bold \textbf{Property(F)} will be evaluated to false, since it is the last of a big disjunction.
-\subsection{Evaluation algorithm}
+\subsection{Evaluation algorithm}\label{algo}
 As already mentioned, monitored formulas ($\underline{\text{mon}}$) evolve at every event, to store relevant informations about past istants.
@@
 By combining the definitions for the future and past time LTLs defined above, we obtain a temporal logic over the future, present and past, in which one can freely intermix the future and past time modalities.
+
 \section{MONID: Structure analysis}
 MONID is a prototype which can detect intrusions either online or offline.
-System-level event data is delivered to MONID server either directly from relevant log files (offline) or online via application code that has been properly instrumented; to create a single event trace, the server combines events from numerous sources according to timestamps and preprocesses them into an abstract intermediate form.
+System-level event data is delivered to MONID server either directly from relevant log files (offline) or online via application code that has been properly instrumented.
+In the online mode, MONID runs as a server that receives streams of events
+from various sources and to generate these events, the different logging modules are instrumented so that they filter and send relevant events to the server.
+To create a single event trace, the server combines events from numerous sources according to timestamps and preprocesses them into an abstract intermediate form.
 Note that sometimes it is necessary to watch events from multiple sources in order to detect certain attacks; then, MONID monitor checks this event trail against a predetermined criteria, and if the specification is broken, an intrusion alarm is raised.
-The entire framework is represented in figure~.
+While, in the offline mode, MONID reads various log files and sends an event corresponding each log entry to the server.
+The server then processes the event stream as before to detect intrusion.
 
-\section{Example of attacks}
+The entire framework is represented in Figure~\ref{monid}.
+In paper~\cite{naldurg2004temporal}, the implementation was written in Java and available as a library; there were two available methods used.
+The first one is \textit{parse} which takes a file containing a list of EAGLE specifications and monitor formulas and compiles them into a data structure used to represent the \textit{monitor} represented in Figure~\ref{monid}.
+The client software iteratively invokes the method \textit{evaluate} for each event after compilation; this call internally adjusts the monitor formulas in accordance with Subsection~\ref{algo}.
+An error notice is produced or a pre-specified mechanism is triggered to take a corrective action if the monitored formulas become false at any stage.
+\begin{figure}[]
+	\centering
+	\includegraphics[width=\textwidth]{images/monid.png}
+	\caption{MONID Architecture.}
+	\label{monid}
+\end{figure}
+
+Some evaluations were also have been done in~\cite{naldurg2004temporal}, using the DARPA Intrusion Detection Evaluation dataset~\cite{mchugh2000testing}.
+The experiment was done using offline mode, against some of the attacks presented in Section~\ref{attacks}.
+Although the used hardware is now outdated and obsolete, the overhead was rather modest, thus this method should still be helpful in theory today.
+
+\section{Example of attacks}\label{attacks}
 

### 2023-09-06 Paper/main.tex
update

@@
 \begin{align*}
-& \underline{\text{min}}\ \text{Next}(\underline{\text{Form}}\ F) = \bigcirc F \\
-& \underline{\text{max}}\ \text{Always}(\underline{\text{Form}}\ F) = F \land \bigcirc \text{Always}(F) \\
-& \underline{\text{min}}\ \text{Eventually}(\underline{\text{Form}}\ F) = F \lor \bigcirc \text{Eventually}(F) \\
-& \underline{\text{min}}\ \text{Until}(\underline{\text{Form}}\ F_1, \underline{\text{Form}}\ F_2) = F_2 \lor (F_1 \land \bigcirc \text{Until}(F_1,F_2))
+& \underline{\text{min}}\ \mathtt{Next}(\underline{\text{Form}}\ F) = \bigcirc F \\
+& \underline{\text{max}}\ \mathtt{Always}(\underline{\text{Form}}\ F) = F \land \bigcirc \mathtt{Always}(F) \\
+& \underline{\text{min}}\ \mathtt{Eventually}(\underline{\text{Form}}\ F) = F \lor \bigcirc \mathtt{Eventually}(F) \\
+& \underline{\text{min}}\ \mathtt{Until}(\underline{\text{Form}}\ F_1, \underline{\text{Form}}\ F_2) = F_2 \lor (F_1 \land \bigcirc \mathtt{Until}(F_1,F_2))
 \end{align*}
 
-As we can see, rules are parameterized by an EAGLE formula (of type $\underline{\text{Form}}$), which means that we will be able to write EAGLE formulas such as $\text{Always}(\text{Eventually}(x>0))$.
-Also, the $\text{Always}$ operator is defined as maximal solution of the equation $X=F\land \bigcirc X$, while the $\text{Eventually}$ operator represents the minimal solution to the equation $X=F\lor \bigcirc X$.
+As we can see, rules are parameterized by an EAGLE formula (of type $\underline{\text{Form}}$), which means that we will be able to write EAGLE formulas such as $\mathtt{Always}(\mathtt{Eventually}(x>0))$.
+Also, the $\mathtt{Always}$ operator is defined as maximal solution of the equation $X=F\land \bigcirc X$, while the $\mathtt{Eventually}$ operator represents the minimal solution to the equation $X=F\lor \bigcirc X$.
 
@@
 \begin{align*}
-    & \square{(\text{(action = login)}\rightarrow \underline{\text{let}}\ \text{k} =userId\ \underline{\text{in}}\ \lozenge(\text{action = logout}\land userId = k))}
+    & \square{((action = \mathtt{login})\rightarrow \underline{\text{let}}\ k=userId\ \underline{\text{in}}\ \lozenge(action = \mathtt{logout}\land userId = k))}
 \end{align*}
-In this formula we use the operator \underline{let} \_ \underline{in} \_ to bind the value of $userId$ in the current event to the local variable $k$ whenever $\text{(action = login)}$ in the current event; we then impose the condition that the value of $userId$ in some event in future must be same as the user id bound to k and that the action of the event must be logout.
+In this formula we use the operator \underline{let} \_ \underline{in} \_ to bind the value of $userId$ in the current event to the local variable $k$ whenever $(action = \mathtt{login})$ in the current event; we then impose the condition that the value of $userId$ in some event in future must be same as the user id bound to k and that the action of the event must be logout.
 In EAGLE we can express the same property with the following rules:
 \begin{align*}
-    & \underline{\text{min}}\ \text{EvLogout}(\underline{\text{string}}\ k) = (\text{action = logout}\land userId = k) \lor \bigcirc \text{EvLogout}(k) \\
-    & \underline{\text{mon}}\ M_2 = ((\text{action = login})\rightarrow \text{EvLogout}(userId)) 
+    & \underline{\text{min}}\ \mathtt{EvLogout}(\underline{\text{string}}\ k) = (action = \mathtt{logout}\land userId = k) \lor \bigcirc \mathtt{EvLogout}(k) \\
+    & \underline{\text{mon}}\ M_2 = ((action = \mathtt{login})\rightarrow \mathtt{EvLogout}(userId)) 
 \end{align*}
@@
     \begin{align*}
-        & \text{LoginLogoutEvent}\{userId:\underline{\text{string}},action: \underline{\text{int}}, time: \underline{\text{double}}\} 
+        & \mathtt{LoginLogoutEvent}\{userId:\underline{\text{string}},\ action: \underline{\text{int}},\ time: \underline{\text{double}}\} 
     \end{align*}
-    An example of event could be: $\{userId:\text{"Lori"},action:\text{login},time:\text{20}\}$
+    An example of event could be: $\{userId:\mathtt{"Lori"},\ action:\mathtt{login},\ time:\mathtt{20}\}$
     \item For each attack, there is a formula $F$ which specifies the absence of it.
@@
 \begin{align*}
-    & \underline{\text{min}}\ \text{Property}(\underline{\text{Form}}\ F) =F \lor \bigcirc \text{Property}(F)
+    & \underline{\text{min}}\ \mathtt{Property}(\underline{\text{Form}}\ F) =F \lor \bigcirc \mathtt{Property}(F)
 \end{align*}
@@
 \begin{align*}
-    & evaluate(\ldots evaluate(evaluate(\text{Property},\alpha_1),\alpha_2),\ldots,\alpha_n)=true
+    & evaluate(\ldots evaluate(evaluate(\mathtt{Property},\alpha_1),\alpha_2),\ldots,\alpha_n)=true
 \end{align*}
@@
 \begin{align*}
-    & F \lor \bigcirc(F \lor \bigcirc(\ldots F \lor \bigcirc \textbf{Property(F)})) 
+    & F \lor \bigcirc(F \lor \bigcirc(\ldots F \lor \bigcirc \mathbf{Property(F)})) 
 \end{align*}
-and the bold \textbf{Property(F)} will be evaluated to false, since it is the last of a big disjunction.
+and the bold $\mathbf{Property(F)}$ will be evaluated to false, since it is the last of a big disjunction.
 \subsection{Evaluation algorithm}\label{algo}
@@
 The evolved formula's value is determined at the conclusion of the event series; if it is true, the formula is satisfied by the event sequence; otherwise, it is violated.
-Formally, a formula $F' = \text evaluate(F, \alpha_i)$ is created when a formula $F$ is evaluated at an event $\alpha_i = \sigma(i)$ and has the condition that $\sigma, i \models F$ if and only if $\sigma, i+1 \models F'$.
+Formally, a formula $F' = evaluate(F, \alpha_i)$ is created when a formula $F$ is evaluated at an event $\alpha_i = \sigma(i)$ and has the condition that $\sigma, i \models F$ if and only if $\sigma, i+1 \models F'$.
 With $F$ being the evolved formula, we compute the boolean function $value(F)$ at the conclusion of the trace.
@@
     %\begin{align*}
-        \text{value}(\underline{\text{true}}) & = \underline{\text{true}} \\
-        \text{value}(\underline{\text{false}}) & = \underline{\text{false}} \\ 
-        \text{value}(\text{exp}) & = \underline{\text{false}} \\
-        \text{value}(F_1\ \text{op}\ F_2) & = \text{value}(F_1)\ \text{op}\ \text{value}(F_2) \\
-        \text{value}(\neg{F}) & = \neg{\text{value}(F)}  \\
-        \text{value}(\bigcirc F) & =
+        value(\underline{\text{true}}) & = \underline{\text{true}} \\
+        value(\underline{\text{false}}) & = \underline{\text{false}} \\ 
+        value(\text{exp}) & = \underline{\text{false}} \\
+        value(F_1\ \text{op}\ F_2) & = value(F_1)\ \text{op}\ value(F_2) \\
+        value(\neg{F}) & = \neg{value(F)}  \\
+        value(\bigcirc F) & =
         \begin{cases}
@@
         \end{cases}\\
-        \text{value}(F_1 \cdot F_2,\alpha_i) & = \text{value}(F_1)\land \text{value}(F_2) \\ \\
+        value(F_1 \cdot F_2,\alpha_i) & = value(F_1)\land value(F_2) \\ \\
    % \end{align*}
@@
     %\begin{align*}
-        \text{update}(\underline{\text{true}},\alpha_i) & = \underline{\text{true}} \\
-        \text{update}(\underline{\text{false}},\alpha_i) & = \underline{\text{false}} \\ 
-        \text{update}(\text{exp},\alpha_i) & = \text{exp}\\
-        \text{update}(F_1\ \text{op}\ F_2,\alpha_i) & = \text{update}(F_1,\alpha_i)\ \text{op}\ \text{update}(F_2,\alpha_i) \\
-        \text{update}(\neg{F},\alpha_i) & = \neg{\text{update}(F,\alpha_i)}  \\
-        \text{update}(\bigcirc F,\alpha_i) & = \text{update}(F,\alpha_i)  \\
-        \text{update}(F_1 \cdot F_2,\alpha_i) & = \text{update}(F_1,\alpha_i)\cdot F_2
+        update(\underline{\text{true}},\alpha_i) & = \underline{\text{true}} \\
+        update(\underline{\text{false}},\alpha_i) & = \underline{\text{false}} \\ 
+        update(\text{exp},\alpha_i) & = \text{exp}\\
+        update(F_1\ \text{op}\ F_2,\alpha_i) & = update(F_1,\alpha_i)\ \text{op}\ update(F_2,\alpha_i) \\
+        update(\neg{F},\alpha_i) & = \neg{update(F,\alpha_i)}  \\
+        update(\bigcirc F,\alpha_i) & = update(F,\alpha_i)  \\
+        update(F_1 \cdot F_2,\alpha_i) & = update(F_1,\alpha_i)\cdot F_2
     \end{align*}
@@
 \begin{align*}
-    &\underline{\text{max}}\ \text{Always}(\underline{\text{Form}}\ F)  = F \land \bigcirc \text{Always}(F) \\
-    &\underline{\text{min}}\ \text{EvTimedLogout}(\underline{\text{string }} k, \underline{\text{double }} t, \underline{\text{double }} \delta) = (\text{time} - t \leq \delta) \\
-    &\quad\quad\quad\quad \land ((action = \text{logout} \land userId = k) \lor \bigcirc \text{EvTimedLogout}(k, t, \delta)) \\
-    &\underline{\text{mon}}\ M_3  = \text{Always}((action = \text{login}) \rightarrow \text{EvTimedLogout}(userId,time, 100))
+    &\underline{\text{max}}\ \mathtt{Always}(\underline{\text{Form}}\ F)  = F \land \bigcirc \mathtt{Always}(F) \\
+    &\underline{\text{min}}\ \mathtt{EvTimedLogout}(\underline{\text{string }} k, \underline{\text{double }} t, \underline{\text{double }} \delta) = (\text{time} - t \leq \delta) \\
+    &\quad\quad\quad\quad \land ((action = \mathtt{logout} \land userId = k) \lor \bigcirc \mathtt{EvTimedLogout}(k, t, \delta)) \\
+    &\underline{\text{mon}}\ M_3  = \mathtt{Always}((action = \mathtt{login}) \rightarrow \mathtt{EvTimedLogout}(userId,time, 100))
 \end{align*}
-and a trace $\sigma=\alpha_1,\alpha_2$, where $\alpha_1=\{userId:\text{"Lori"},action:\text{login},time:\text{17.0}\}$ and $\alpha_1=\{userId:\text{"Lori"},action:\text{logout},time:\text{150.0}\}$, the initial formula $M_3$ gets modified.
+and a trace $\sigma=\alpha_1,\alpha_2$, where $\alpha_1=\{userId:\mathtt{"Lori"},\ action:\mathtt{login},\ time:\mathtt{17.0}\}$ and $\alpha_1=\{userId:\mathtt{"Lori"},\ action:\mathtt{logout},\ time:\mathtt{150.0}\}$, the initial formula $M_3$ gets modified.
 At time $1$, when $\alpha_1$ is analyzed, $M_3$ changes as follows:
 \begin{align*}
-    &evaluate(M_3,\alpha_1)  = \text{EvTimedLogout("Lori",17.0, 100)} \\ 
-    &\quad\quad\quad\quad \land\text{Always}((action = \text{login}) \rightarrow \text{EvTimedLogout}(userId,time, 100))
+    &evaluate(M_3,\alpha_1)  = \mathtt{EvTimedLogout("Lori",17.0, 100)} \\ 
+    &\quad\quad\quad\quad \land\mathtt{Always}((action = \mathtt{login}) \rightarrow \mathtt{EvTimedLogout}(userId,time, 100))
 \end{align*}
@@
 \begin{align*}
-    & \underline{\text{min}}\ \text{Previous}(\underline{\text{Form}}\ F) = \odot F \\
-    & \underline{\text{max}}\ \text{AlwaysInPast}(\underline{\text{Form}}\ F) = F \land \odot \text{AlwaysInPast}(F) \\
-    & \underline{\text{min}}\ \text{EventuallyInPast}(\underline{\text{Form}}\ F) = F \lor \odot \text{EventuallyInPast}(F) \\
-    & \underline{\text{min}}\ \text{Since}(\underline{\text{Form}}\ F_1, \underline{\text{Form}}\ F_2) = F_2 \lor (F_1 \land \odot \text{Since}(F_1,F_2))
+    & \underline{\text{min}}\ \mathtt{Previous}(\underline{\text{Form}}\ F) = \odot F \\
+    & \underline{\text{max}}\ \mathtt{AlwaysInPast}(\underline{\text{Form}}\ F) = F \land \odot \mathtt{AlwaysInPast}(F) \\
+    & \underline{\text{min}}\ \mathtt{EventuallyInPast}(\underline{\text{Form}}\ F) = F \lor \odot \mathtt{EventuallyInPast}(F) \\
+    & \underline{\text{min}}\ \mathtt{Since}(\underline{\text{Form}}\ F_1, \underline{\text{Form}}\ F_2) = F_2 \lor (F_1 \land \odot \mathtt{Since}(F_1,F_2))
 \end{align*}
@@
 The experiment was done using offline mode, against some of the attacks presented in Section~\ref{attacks}.
-Although the used hardware is now outdated and obsolete, the overhead was rather modest, thus this method should still be helpful in theory today.
+Although the used hardware is now outdated and obsolete, the overhead was rather modest, thus this method should still be theoretically helpful today.
 
 \section{Example of attacks}\label{attacks}
+We give some examples of how EAGLE can be utilized to build formulas that correlate to desired execution trace characteristics of a system under observation in this section.
+In this context, a trace $\sigma$ that violates this specification is referred to as an intrusion or an attack; to explain how the framework may be used, we use examples from actual attacks illustrated in~\cite{naldurg2004temporal}.
+These illustrations demonstrate the expressive power of EAGLE formalism as well as its numerous features.
+In all the following examples, the rule Always is the one defined in Section~\ref{sec:ltl}.
+
+\textbf{Smurf attacks.}
+A smurf attack~\cite{zargar2009identification} is a type of Distributed Denial of Service (DDoS) attack that targets computer networks; in this scenario, the attacker sends a large volume of Internet Control Message Protocol (ICMP) echo request packets (commonly known as "ping" packets) to a network's broadcast address.
+This broadcast address is the address that reaches all devices on the targeted network.
+The attacker spoofs the source IP address of these ICMP packets to make it appear as if they are originating from the victim's IP address and since the broadcast address sends the request to all devices on the network, all devices will respond with ICMP echo replies to the victim's IP address.
+This results in a flood of traffic overwhelming the victim's network and causing it to become unreachable or severely degraded in performance.
+To detect this kind of attacks, we need to work on network logs; we will use a network tool called \textit{tcpdump}.
+The monitored formula that describes the absence of the attack is the following:
+\begin{align*}
+    & \underline{\text{max}}\ \mathtt{smurfAttack()} = (type = \text{"ICMP"})\land isBroadcast(ip) \\
+    & \underline{\text{mon}}\ \mathtt{smurfSafety} =\ \mathtt{Always}(\neg\ \mathtt{smurfAttack()}) 
+\end{align*}
+as we can see, the \textit{type} field of a record schema describes the type of network packet, while the \textit{ip} field is the destination of the packet.The $\mathtt{smurfAttack}$ rule determines whether the packet is a ping packet and whether the source's IP address is a broadcast ip, i.e.
+all the host bits are set to 1.
+Thus, \textit{"Always there is no attack"} can be used to describe how the system behaves well in response to this attack.
+
+\textbf{Cookie-Stealing attacks.}
 

### 2023-09-06 Paper/main.tex
update

@@
 \begin{align*}
-    & \underline{\text{max}}\ \mathtt{smurfAttack()} = (type = \text{"ICMP"})\land isBroadcast(ip) \\
-    & \underline{\text{mon}}\ \mathtt{smurfSafety} =\ \mathtt{Always}(\neg\ \mathtt{smurfAttack()}) 
+    & \underline{\text{max}}\ \mathtt{SmurfAttack()} = (type = \text{"ICMP"})\land isBroadcast(ip) \\
+    & \underline{\text{mon}}\ \mathtt{SmurfSafety} =\ \mathtt{Always}(\neg\ \mathtt{SmurfAttack()}) 
 \end{align*}
-as we can see, the \textit{type} field of a record schema describes the type of network packet, while the \textit{ip} field is the destination of the packet.The $\mathtt{smurfAttack}$ rule determines whether the packet is a ping packet and whether the source's IP address is a broadcast ip, i.e.
+as we can see, the \textit{type} field of a record schema describes the type of network packet, while the \textit{ip} field is the destination of the packet.The $\mathtt{SmurfAttack}$ rule determines whether the packet is a ping packet and whether the source's IP address is a broadcast ip, i.e.
 all the host bits are set to 1.
@@
 
-\textbf{Cookie-Stealing attacks.}
+\textbf{Cookie-stealing attacks.}
+A cookie is a technique for tracking sessions that a web application server sends to a web client and uses to keep track of client-specific session data.
+These cookies that serve as authentication tokens are automatically included in requests by clients and in this scenario we assume that a session is exclusively identified by the client's IP address.
+When a malicious user uses an old cookie provided to a different IP address to take over a session, it is considered an attack.In order to monitor this attack we need to look at a web-server (application-level) log that contains a record of all sessions that the server participates in, along with session-specific state information.
+According to the formula below, a specific cookie must always be utilized by the same IP address.
+%%  esempio sbagliato IMO
+%\begin{align*}
+%    & \underline{\text{min}}\ \mathtt{SafeUse}(\underline{\text{string}}\ c,\underline{\text{int}}\ i) = ((name=c)\rightarrow(ip=i))\land\odot\ \mathtt{SafeUse}(c,i) \\
+%    & \underline{\text{mon}}\ \mathtt{CookieSafe} =\ \mathtt{Always}(\mathtt{SafeUse}(name,ip)) 
+%\end{align*}
+\begin{align*}
+    & \underline{\text{min}}\ \mathtt{Hijack}(\underline{\text{string}}\ c,\underline{\text{int}}\ i) = ((name=c)\land\neg(ip=i))\lor\odot\ \mathtt{Hijack}(c,i) \\
+    & \underline{\text{mon}}\ \mathtt{CookieSafe} =\ \mathtt{Always}(\neg\mathtt{Hijack}(name,ip)) 
+\end{align*}
+A trace that violates this formula therefore encodes a cookie-stealing attack.
+The rule $\mathtt{Hijack}$ checks wheter it occurs a situation in which we have the same coookie $c$ associated with two different ip $i$; if it's true then it means that a cookie has been stolen and reused by an attacker.
+With $\mathtt{CookieSafe}$ we want to check that it never occurs a situation just as the one described.
 

### 2023-09-06 Paper/main.tex
update

@@
 
+\textbf{Multi-domain Buffer Overflows}
+According to~\cite{piromsopa2006buffer}, a buffer-overflow attack is an attack that (possibly implicitly) uses memory-manipulating operations to overflow a buffer which results in the modification of an address to point to malicious or unexpected code.
+In this case, we will integrate data from various log sources; to be more exact, we will examine both network and web server access logs to determine when a server buffer overflow attack has been performed.
+We search through network packets for binary data; then web server's access logs are examined to see if a corresponding event, in which the web server correctly terminates the connection after receiving some binary data, can be located.
+If no matching log record is discovered within a predetermined timeout, it means that the buffer overflow attack was successful.
+The following formula captures this scenario:
+
 \section{Conclusions}

### 2023-09-06 Paper/main.tex
update

@@
 \begin{align*}
-    & \underline{\text{min}}\ \mathtt{Hijack}(\underline{\text{string}}\ c,\underline{\text{int}}\ i) = ((name=c)\land\neg(ip=i))\lor\odot\ \mathtt{Hijack}(c,i) \\
+    & \underline{\text{min}}\ \mathtt{Hijack}(\underline{\text{string}}\ c,\underline{\text{string}}\ i) = ((name=c)\land\neg(ip=i))\lor\odot\ \mathtt{Hijack}(c,i) \\
     & \underline{\text{mon}}\ \mathtt{CookieSafe} =\ \mathtt{Always}(\neg\mathtt{Hijack}(name,ip)) 
@@
 
-\textbf{Multi-domain Buffer Overflows}
+\textbf{Multi-domain Buffer Overflows attacks.}
 According to~\cite{piromsopa2006buffer}, a buffer-overflow attack is an attack that (possibly implicitly) uses memory-manipulating operations to overflow a buffer which results in the modification of an address to point to malicious or unexpected code.
@@
 The following formula captures this scenario:
+\begin{align*}
+    & \underline{\text{min}}\ \mathtt{EventuallyClosed}(\underline{\text{long}}\ t,\underline{\text{long}}\ d,\underline{\text{string}}\ i1,\underline{\text{string}}\ i2) = (time-t<d)\\
+    & \quad\quad\land ((ip1=i1\land ip2=i2\land log=web\land type=closed)\lor \bigcirc\mathtt{EventuallyClosed}(t,d,i1,i2) ) \\
+    & \underline{\text{mon}}\ \mathtt{BufferSafe} =\ \mathtt{Always}((log=network\land type=binary) \\
+    & \quad\quad\rightarrow\mathtt{EventuallyClosed}(time,100,ip1,ip2))
+\end{align*}
+In each record we have the $log$ field which indicates the source of the log, the $time$ represents the istant in which the event occurs, $ip1$ represents the source address, $ip2$ represents the destination IP address and finally the $type$ field can be $binary, closed,\ldots$ We assume that a connection is identified by the tuple $(ip1,ip2)$.
+In order, the $\mathtt{EventuallyClosed}$ rule states that the connection between IP addresses $i1$ and $i2$ must end eventually within time $d$; finally, the monitor rule $\mathtt{BufferSafe}$ states that it is always the case that if there is an event of $binary$ access in the network log then eventually within time $100$ there must be a matching event in the webserver log that denotes the closing of the connection.
 
+\textbf{Password guessing attacks.}
 \section{Conclusions}

### 2023-09-07 Paper/main.tex
update

@@
 If no matching log record is discovered within a predetermined timeout, it means that the buffer overflow attack was successful.
-The following formula captures this scenario:
+The following formulas captures this scenario:
 \begin{align*}
@@
 \textbf{Password guessing attacks.}
+Assuming the system accepts telnet connections to remote login; it won't take long until the password is compromised if an attacker is permitted to guess any amount of passwords for a specific user name.
+If a user guesses an incorrect password more than three times in a short amount of time, most systems end the telnet session.
+To prevent an attacker from succeeding by starting numerous temporary sessions, some systems cap the total number of unsuccessful login attempts over a longer time frame; in this case we need to have access to the host's audit logs.
+The formulas that capture this scenario are the following:
+\begin{align*}
+    & \underline{\text{max}}\ \mathtt{Failure}() = (type=login)\land \neg success\\
+    & \underline{\text{max}}\ \mathtt{Guess}(\underline{\text{string}}\ i) = (ip=i)\land \mathtt{Failure}()\\
+    & \underline{\text{max}}\ \mathtt{Counter}(\underline{\text{long}}\ t, \underline{\text{long}}\ d, \underline{\text{int}}\ c,\underline{\text{string}}\ i,\underline{\text{int}}\ C) = (time-t<d)\\
+    & \quad\quad\rightarrow((\mathtt{Guess}(i)\rightarrow(c\leq C\land \bigcirc\mathtt{Counter}(t,d,c+1,i,C)))  \\
+    & \quad\quad\land(\neg \mathtt{Guess}(i)\rightarrow\mathtt{Counter}(t,d,c,i,C))) \\
+    & \underline{\text{mon}}\ \mathtt{PassGuessSafe} =\ \mathtt{Always}(\mathtt{Failure}()\rightarrow\mathtt{Counter}(time,300,1,ip,3)) 
+\end{align*}
+as we can see, the rules here are numerous.
+In $\mathtt{Counter}$, the arguments $t, d, c, i, C$ represent the rule invocation time, the timeout period, the current number of unsuccessful-guesses count, the source IP address doing the guess, and the threshold count.
+When the number of guesses from the IP address $i$ exceeds the threshold count $C$ within the timeout period $d$, an attack has taken place.
+The parameterized rule $\mathtt{Counter}$ starts with the initial count $c$ set to 1 whenever there is a $\mathtt{Failure}$ during login.
+The number of guesses is increased by one for each instance of an event signifying a failed login attempt from the same IP address within the timeout period $d$.
+The rule $\mathtt{Counter}$ also determines whether the local counter $c$ surpasses the threshold $C$ within time d, and if it does, the entire rule become false, indicating an attack.
+Finally, with the rule $\mathtt{PassGuessSafe}$ we assure that whenever there is a failed login from a specific IP address, we make sure that within time $300$ seconds the number of login-failures from the same IP address must be less than or equal to $3$.
+
 \section{Conclusions}

### 2023-09-07 Paper/main.tex
update

@@
 
+\textbf{Port Scanning attacks.}
+According to~\cite{kanlayasiri2000rule}, Port scanning attack is a method for discovering exploitable communication channels that has been used for a long time.
+The key idea is to probe the network ports and then keep the information about them that are useful for an attack.
+Hackers frequently employ a port scan approach to find gaps or weak spots in a network.
+Attackers can use a port scan attack to identify open ports and determine if they are accepting or rejecting data; additionally, it can show whether a company uses firewalls or other active security measures.
+Assuming that our server has multiple network inferfaces, thus multiple IP addresses; the victim machine can presume that port scans are malicious when the counter of reached ports overcomes a specific threshold in a short amount of time.
+The formulas are the following:
+\begin{align*}
+    & \underline{\text{max}}\ \mathtt{NewPort}(\underline{\text{string}}\ i1,\underline{\text{string}}\ i2,\underline{\text{Set}}\ S) = (i1=ip1)\land(i2=ip2)\land (port \notin S)\\
+    & \underline{\text{max}}\ \mathtt{Counter}(\underline{\text{long}}\ t, \underline{\text{long}}\ d, \underline{\text{int}}\ c,\underline{\text{string}}\ i1,\underline{\text{string}}\ i2,\underline{\text{Set}}\ S,\underline{\text{int}}\ C) = (time-t<d)\\
+    & \quad\quad\rightarrow((\mathtt{NewPort}(i1,i2,S)\rightarrow(c\leq C\land \bigcirc\mathtt{Counter}(t,d,c+1,i1,i2,S\cup\{port\},C)))  \\
+    & \quad\quad\land(\neg \mathtt{NewPort}(i1,i2,S)\rightarrow\mathtt{Counter}(t,d,c,i1,i2,S,C))) \\
+    & \underline{\text{mon}}\ \mathtt{PortScanSafe} =\ \mathtt{Always}(\mathtt{Counter}(time,100,1,ip1,ip2,\{port\},10)) 
+\end{align*}
+in this case we have two IPs; $ip1$ is the source IP, while $ip2$ is the destination one.
+The parameterized $\mathtt{Counter}$ rule states that the number of port scans between a source and destination IP  address pair $(i1,i2)$ recorded do not surpass a threshold $C$ within time $d$.
+The rule $\mathtt{NewPort}$ checks if the port number,
+involved in any communication between the IP addresses $i1$ and $i2$, exists in the set $S$ of all port numbers (involved in all communications between $i1$ and $i2$) within the timeout period $d$.
+
 \section{Conclusions}

### 2023-09-07 Paper/main.tex
update

@@
 rule application, for which get generated based on the definition of rules in the specification.
-The definitions of these functions on the different primitive operators are the following ($op\text{ can be }\lor , \land, \rightarrow $):
+The definitions of these functions on the different primitive operators (except the previous operator $\odot$) are the following ($op\text{ can be }\lor , \land, \rightarrow $):
 %evaluate
@@
         \end{cases}\\
+        value(F_1 \cdot F_2,\alpha_i) & = value(F_1)\land value(F_2) \\ 
+        value(\mathtt{R}(F_1,\ldots,F_n)) & =
+        \begin{cases}
+            \underline{\text{true}} & \text{if }\mathtt{R}\text{ is } \underline{\text{max}}\\
+            \underline{\text{false}} & \text{if }\mathtt{R}\text{ is } \underline{\text{min}}\\
+        \end{cases}\\
         value(F_1 \cdot F_2,\alpha_i) & = value(F_1)\land value(F_2) \\ \\
@@
     }
+Focusing of \textit{Future Time Operators}, we have:
+\begin{align*}
+    evaluate(\mathtt{Next}(F),s)&=evaluate(\bigcirc F,s) \\
+    evaluate(\mathtt{Always}(F),s)&=evaluate(F\land\bigcirc\mathtt{Always}(F),s) \\
+    evaluate(\mathtt{Eventually}(F),s)&=evaluate(F\lor\bigcirc\mathtt{Eventually}(F),s) \\
+    evaluate(\mathtt{Until}(F_1,F_2),s)&=evaluate(F_2\lor(F_1\land\bigcirc\mathtt{Until}(F_1,F_2)),s) \\ \\
+    update(\mathtt{Next}(F),s)&=\mathtt{Next}(update(F,s)) \\
+    update(\mathtt{Always}(F),s)&=\mathtt{Always}(update(F,s)) \\
+    update(\mathtt{Eventually}(F),s)&=\mathtt{Eventually}(update(F,s)) \\
+    update(\mathtt{Until}(F_1,F_2),s)&=\mathtt{Until}(update(F_1,s),update(F_2,s))
+\end{align*}
+while, for the \textit{Past Time Operators} the $\odot$ operator is removed and instead a rule $\mathtt{Previous}$ is introduced; more details can be found in~\cite{barringer2004program}.
+
 %\end{minipage}

### 2023-09-08 Paper/main.tex
update

@@
 The definition of the function $evaluate: \underline{\text{Form}} \times \underline{\text{State}} \to \underline{\text{Form}}$ uses an auxiliary function $update: \underline{\text{Form}} \times \underline{\text{State}} \to \underline{\text{Form}}$ to pre-evaluate a formula if it is guarded by the previous operator $\bigodot$.
+Note that $\underline{\text{Form}}$ in the practical use represents the monitored formula, while $\underline{\text{State}}$ represents the current event.
 At the end (or at the beginning) of a trace, the
@@
             \underline{\text{false}} & \text{if }\mathtt{R}\text{ is } \underline{\text{min}}\\
-        \end{cases}\\
-        value(F_1 \cdot F_2,\alpha_i) & = value(F_1)\land value(F_2) \\ \\
+        \end{cases}\\ \\
            % \end{align*}
@@
 \end{align*}
-and a trace $\sigma=\alpha_1,\alpha_2$, where $\alpha_1=\{userId:\mathtt{"Lori"},\ action:\mathtt{login},\ time:\mathtt{17.0}\}$ and $\alpha_1=\{userId:\mathtt{"Lori"},\ action:\mathtt{logout},\ time:\mathtt{150.0}\}$, the initial formula $M_3$ gets modified.
+and a trace $\sigma=\alpha_1,\alpha_2$, where $\alpha_1=\{userId:\mathtt{"Lori"},\ action:\mathtt{login},\ time:\mathtt{17.0}\}$ and $\alpha_2=\{userId:\mathtt{"Lori"},\ action:\mathtt{logout},\ time:\mathtt{150.0}\}$, the initial formula $M_3$ gets modified.
 At time $1$, when $\alpha_1$ is analyzed, $M_3$ changes as follows:
@@
 % Note that UNLESS is maximal
-By combining the definitions for the future and past time LTLs defined above, we obtain a temporal logic over the future, present and past, in which one can freely intermix the future and past time modalities.
+By combining the definitions for the future and past time LTLs defined above, we obtain a temporal logic over the future, present and past, in which one can freely intermix the future and past time modalities; thus, we can state that EAGLE is expressive at least as LTL+P.
+% a differenza di ltl non salva tutta la sequenza ma solo una formula che riassume tutto. La seconda componente non è uno stato (insieme di cose vere) ma un record
+% EAGLE è almeno espressivo quanto LTLP
 

### 2023-09-08 Paper/main.tex
update

@@
 %% INTRODURRE VELOCE ESEMPIO E STRUTTURA DEL PROGRAMMA
-
 %%%%%   DA RISCRIVERE ALLA FINE
-In Section 2 we will analyze the basics of EAGLE.
-In Section 3 we will analyze MONID framework and some evalutations on the performances.
-In Section 4 we will see some attacks specifications.
-In section 5 we will conclude.
+In Section 2, we'll deeply explore EAGLE, covering everything from its syntax to semantics and the evaluation algorithm, with the add of practical examples.
+Moving on to Section 3, we'll delve into the MONID framework, which employs EAGLE to construct an intrusion detection system.
+Section 4 will focus on specific EAGLE rules tailored for the detection of various types of attacks and finally, in Section 5, we will draw our conclusions.
 %%%%5
@@
 
-It's essential to emphasize that the logic EAGLE operates with a \textit{finite} trace semantics; the monitoring algorithm, as written before, can check whetet the formula is satisfied or not only at the conclusion of a trace.
+It's essential to emphasize that the logic EAGLE operates with a \textit{finite} trace semantics; the monitoring algorithm, as written before, can check whether the formula is satisfied or not only at the conclusion of a trace.
 However, in intrusion detection scenarios, the notion of the "end of a trace" makes no sense since event sequences can theoretically be infinite.
@@
 \section{Conclusions}
+In conclusion, we have introduced a new temporal logic named EAGLE, which serves as the foundation for the MONID framework, which is a useful tool for system intrusion detection.
+MONID continuously monitor these attack specifications during system execution and when the observed execution contradicts the monitored formula, we detect an intrusion.
+We have showcased this by creating formulas to identify several common attacks and testing our monitoring algorithm on large event logs provided by DARPA.
+These examples can serve as templates for specifying many other types of attacks.
+The obtained results were excellent, both in terms of detected attacks and system overhead.
 

### 2023-09-09 Paper/main.tex
update

@@
 \end{align*}
-% Note that UNLESS is maximal
 By combining the definitions for the future and past time LTLs defined above, we obtain a temporal logic over the future, present and past, in which one can freely intermix the future and past time modalities; thus, we can state that EAGLE is expressive at least as LTL+P.
 % a differenza di ltl non salva tutta la sequenza ma solo una formula che riassume tutto. La seconda componente non è uno stato (insieme di cose vere) ma un record
-% EAGLE è almeno espressivo quanto LTLP
 
@@
 According to the formula below, a specific cookie must always be utilized by the same IP address.
-%%  esempio sbagliato IMO
-%\begin{align*}
-%    & \underline{\text{min}}\ \mathtt{SafeUse}(\underline{\text{string}}\ c,\underline{\text{int}}\ i) = ((name=c)\rightarrow(ip=i))\land\odot\ \mathtt{SafeUse}(c,i) \\
-%    & \underline{\text{mon}}\ \mathtt{CookieSafe} =\ \mathtt{Always}(\mathtt{SafeUse}(name,ip)) 
-%\end{align*}
+
 \begin{align*}

### 2023-09-09 Presentation/presentation.tex
update

\documentclass[aspectratio=169,t,xcolor=table]{beamer}
\usepackage[utf8]{inputenc}

\usepackage{booktabs} 
\usepackage{subcaption}
\usepackage{epigraph}
\usepackage[outercaption]{sidecap}    
\usepackage{tikz}

\usetheme{Ufg}

%-------------------------------------theorems--------------
\newtheorem{conj}{Conjetura}
\newtheorem{defi}{Definition}
\newtheorem{teo}{Teorema}
\newtheorem{lema}{Lema}
\newtheorem{prop}{Proposição}
\newtheorem{cor}{Corolário}
\newtheorem{ex}{Example}
\newtheorem{exer}{Exercício}

\setbeamertemplate{theorems}[numbered]
\setbeamertemplate{caption}[numbered]

%-------------------------------------------------------------%
%----------------------- Primary Definitions -----------------%

\definecolor{color1}{RGB}{0,0,90} % Color of the article title and sections
\definecolor{color2}{RGB}{0,20,20} % Color of the boxes behind the abstract and headings
\definecolor{keys1}{rgb}{0.0, 0.29, 0.33}
\definecolor{keys2}{rgb}{0.25, 0.7, 0.55}
\definecolor{keys3}{rgb}{0.1, 0.3, 0.4}
\definecolor{keys4}{rgb}{0.21, 0.46, 0.53}
\definecolor{strings}{rgb}{0.0, 0.47, 0.44}
\definecolor{comments}{rgb}{0.4, 0.4, 0.5}
\definecolor{terminaltext}{rgb}{1,1,1}
\definecolor{terminalbackground}{rgb}{0.2,0.2,0.25}

\usepackage{listings}

\lstset{
  xleftmargin=8pt,
  xrightmargin=0pt,
  framexleftmargin=0pt,
  framexrightmargin=0pt,
  basicstyle={\fontsize{8pt}{10pt}\ttfamily},
  columns=flexible,
  keepspaces=false,
  showstringspaces=false,
  commentstyle= \color{comments},
  stringstyle= \color{strings},
  breaklines=false,
  postbreak=\mbox{$\hookrightarrow$\space}
}

\lstdefinelanguage{Terminal}{
  framexleftmargin=3pt,
  framexrightmargin=3pt,
  framextopmargin=3pt,
  framexbottommargin=3pt,
  backgroundcolor=\color{terminalbackground},
  basicstyle={\fontsize{8pt}{10pt}\ttfamily\color{terminaltext}},
}

%%% Bibliography
\usepackage[style=numeric,backend=biber]{biblatex}
\addbibresource{thud.bib}

% This command set the default Color, is also possible to choose a custom color
\setPrimaryColor{UFGBlue}

% First one is logo in title slide (we recommend use a horizontal image), and second one is the logo used in the remaining slides (we recommend use a square image)
\setLogos{lib/logos/infw.png}{lib/logos/infw2.png}

% -------------------------------------- Title Slide Information
\begin{document}
\title[Inf UFG]{MONID}
\subtitle{A Temporal Logic Based Framework for Intrusion Detection}

\author{Zanolin Lorenzo\inst{1}}

\institute[UFG] % (optional)
{
  \inst{1}%
  DMIF\\
  University of Udine
}
\date{September 2023}
%-----------------------The next statement creates the title page.
\frame[noframenumbering]{\titlepage}

%------------------------------------------------Slide 1
\setLayout{vertical} % This command define the layout. 'vertical' can be replace with 'horizontal', 'blank, 'mainpoint', 'titlepage'

\begin{frame}
    \frametitle{Table of Contents}
    \tableofcontents
\end{frame}
%---------------------------------------------------------

\begin{frame}[allowframebreaks]{References}
    \nocite{*} 
    \printbibliography
\end{frame}

\setLayout{mainpoint}
\setBGColor{DarkGray}
\begin{frame}{}
    \frametitle{Thanks for the attention}
\end{frame}
\setLayout{vertical}

\end{document}

### 2023-09-11 Presentation/presentation.tex
update

@@
 
+\section{Introduction}
+
+\setLayout{mainpoint}
+\setBGColor{DarkPurple}
+\begin{frame}{}
+    \frametitle{Introduction}
+\end{frame}
+\setLayout{vertical}
+
+\begin{frame}
+    \frametitle{What is MONID?}
+    MONID is a prototype which can detect intrusions on a system; in this scenario we are focusing on a \textit{signature-based} approach.\\
+    \vspace{5mm}
+    In order:
+    \begin{enumerate}
+        \item we will use the logic EAGLE to define intrusion patterns using temporal logic formula $\varphi$; in this case the monitored formula will be $\psi=\square (\neg\varphi)$.
+        \item MONID will create a stream of events $\sigma=\alpha_{1},\alpha_{2},\ldots$ obtained from a merge of the logs by ascending time order;
+        \item a monitor will processes each event $\alpha_i$ as it happens and updates the monitored formula $\psi$ to store a relevant summary;
+        \item an intrusion alarm is triggered if, for any reason, $\alpha_{1},\alpha_{2} \ldots\not\models\psi$.
+    \end{enumerate}
+
+\end{frame}
+
+\begin{frame}
+    \frametitle{What is MONID? (cont'd)}
+    The architecture is the following.\\
+    \vspace{5mm}
+    \includegraphics[scale=0.25]{images/monid.png}
+\end{frame}
+
 \begin{frame}[allowframebreaks]{References}

### 2023-09-11 Presentation/presentation.tex
update

@@
 \begin{frame}
+    \frametitle{Intrusion Detection}
+    \textit{Intrusion detection} means maintaining constant surveillance on a system in order to detect any misuse of these weak areas as soon as feasible so that they can be repaired.\\
+    There are three approaches:
+    \begin{itemize}
+        \item \textit{signature-based}: aims to identify patterns and match them with known signs of intrusions;
+        \item \textit{anomaly-based}: can identify new attacks when it detects behavior that differs significantly from previously learned normal behavior;
+        \item \textit{hybrid}: combines the best of both worlds by looking at patterns and one-off events.
+    \end{itemize}
+    \vspace{5mm}
+    We will present MONID which is a \textit{signature-based} intrusion detector.
+\end{frame}
+
+\begin{frame}
     \frametitle{What is MONID?}
-    MONID is a prototype which can detect intrusions on a system; in this scenario we are focusing on a \textit{signature-based} approach.\\
+    MONID is a prototype which can detect intrusions on a system and operates in both online and offline modes.\\
     \vspace{5mm}
@@
     \begin{enumerate}
-        \item we will use the logic EAGLE to define intrusion patterns using temporal logic formula $\varphi$; in this case the monitored formula will be $\psi=\square (\neg\varphi)$.
+        \item we will use the logic \textbf{EAGLE} to define intrusion patterns using temporal logic formula $\varphi$; in this case the monitored formula will be $\psi=\square (\neg\varphi)$.
         \item MONID will create a stream of events $\sigma=\alpha_{1},\alpha_{2},\ldots$ obtained from a merge of the logs by ascending time order;
@@
     The architecture is the following.\\
-    \vspace{5mm}
+    \vspace{2.5mm}
     \includegraphics[scale=0.25]{images/monid.png}
+    \vspace{2.5mm}\\
+    Now, let us start from the basics of EAGLE.
 \end{frame}

### 2023-09-11 Paper/main.tex
update

@@
 \section{EAGLE, a Temporal Monitoring Logic}
-According to~\cite{barringer2004program}, EAGLE offers a succinct but powerful set of primitives, essentially supporting recursive parameterized equations, with a minimal/maximal fix-point semantics together with three temporal operators: next-time ($\bigcirc$), previous-time ($\astrosun$), and concatenation ($\cdot$).
+According to~\cite{barringer2004program}, EAGLE offers a succinct but powerful set of primitives, supporting recursive parameterized equations with a minimal/maximal fix-point semantics together with three temporal operators: next-time ($\bigcirc$), previous-time ($\astrosun$), and concatenation ($\cdot$).
 \subsection{Basics}\label{sec:ltl}

### 2023-09-11 Presentation/presentation.tex
update

@@
     \textit{Intrusion detection} means maintaining constant surveillance on a system in order to detect any misuse of these weak areas as soon as feasible so that they can be repaired.\\
+    \vspace{5mm}
     There are three approaches:
@@
 
-\begin{frame}
-    \frametitle{What is MONID?}
+\begin{frame}[allowframebreaks]{What is MONID?}
     MONID is a prototype which can detect intrusions on a system and operates in both online and offline modes.\\
@@
 
+    The architecture is the following.\\
+    \vspace{2.5mm}
+    \includegraphics[scale=0.25]{images/monid.png}
 \end{frame}
@@
 \begin{frame}
-    \frametitle{What is MONID? (cont'd)}
-    The architecture is the following.\\
+    \frametitle{Assumptions}
+    Two assumptions must be made:
+
+    \vspace{5mm}
+    \begin{enumerate}
+        \item There is a finite sequence of events called $\sigma=\alpha_{1},\ldots ,\alpha_{n}$ that is a merge of the system registered logs organized by ascending time. 
+        The structure of an event record $\alpha_i$ is the following:
+        \begin{align*}
+            & \mathtt{LoginLogoutEvent}\{userId:\underline{\text{string}},\ action: \underline{\text{int}},\ time: \underline{\text{double}}\} 
+        \end{align*}
+        An example of event could be:\\ $\quad\quad\{userId:\mathtt{"Lori"},\ action:\mathtt{login},\ time:\mathtt{20}\}$
+        \item For each attack, there is a formula $\psi$ which specifies the absence of it.
+    \end{enumerate}
+    \vspace{5mm}
+    Now let us start on the basics of EAGLE.
+\end{frame}
+
+\section{EAGLE}
+
+\setLayout{mainpoint}
+\setBGColor{Ocean}
+\begin{frame}{}
+    \frametitle{EAGLE,\\ a Temporal Monitoring Logic}
+\end{frame}
+\setLayout{vertical}
+
+\begin{frame}[allowframebreaks]{Basics}
+    EAGLE offers a succinct but powerful set of primitives, supporting recursive parameterized equations with a minimal/maximal fix-point semantics together with three temporal operators: next-time ($\bigcirc$), previous-time ($\odot$), and concatenation ($\cdot$).\\
     \vspace{2.5mm}
-    \includegraphics[scale=0.25]{images/monid.png}
-    \vspace{2.5mm}\\
-    Now, let us start from the basics of EAGLE.
+    As a result, rules in EAGLE give us the power to create specific temporal operators as well as to bind and modify data.
+This property turns out to be crucial for succinctly expressing executions of attack-safe systems.
+
+    \vspace{2.5mm}
+    EAGLE operates with \textit{finite trace} semantics, meaning it checks formula satisfaction only at the end of a trace.
+However, in intrusion detection where event sequences can be infinite, the goal is to trigger an alarm as soon as a property is violated, thus MONID continuously checks the formula's satisfaction status after each event.
+
+    Let us start with an example; we want to express the property "\textit{Whenever there is a login by any user x, then eventually the user x logs out}".
+In EAGLE we can do it with the following rules:
+    \begin{align*}
+        & \underline{\text{min}}\ \mathtt{EvLogout}(\underline{\text{string}}\ k) = (action = \mathtt{logout}\land userId = k) \lor \bigcirc \mathtt{EvLogout}(k) \\
+        & \underline{\text{mon}}\ M_2 = ((action = \mathtt{login})\rightarrow \mathtt{EvLogout}(userId)) 
+    \end{align*}
+    Once the rules are created, the monitor will evaluate and update the monitored formula $M_2$. 
+    \vspace{2.5mm}
+
+    A possibile trace $\sigma=\alpha_1,\alpha_2$, where $\alpha_1=\{userId:\mathtt{"Lori"},\ action:\mathtt{login},\ time:\mathtt{17.0}\}$ and $\alpha_2=\{userId:\mathtt{"Lori"},\ action:\mathtt{logout},\ time:\mathtt{150.0}\}$ can satisfy $M_2$.
+\end{frame}
+
+\section{Conclusions}
+\setLayout{mainpoint}
+\setBGColor{LightGray}
+\begin{frame}{}
+    \frametitle{Conclusions}
+\end{frame}
+\setLayout{vertical}
+
+\begin{frame}
+    \frametitle{Conclusions}
+
 \end{frame}

### 2023-09-11 Presentation/presentation.tex
update

@@
 In EAGLE we can do it with the following rules:
+    \begin{ex} 
+    \vspace{2.5mm}
+    $ \underline{\text{min}}\ \mathtt{EvLogout}(\underline{\text{string}}\ k) = (action = \mathtt{logout}\land userId = k) \lor \bigcirc \mathtt{EvLogout}(k) $\\
+    $  \underline{\text{mon}}\ M_2 = ((action = \mathtt{login})\rightarrow \mathtt{EvLogout}(userId)) $\\
+    \vspace{2.5mm}
+    Once the rules are created, the monitor will evaluate and update the monitored formula $M_2$. 
+    A possibile trace $\sigma=\alpha_1,\alpha_2 $ , where:\\
+    \vspace{2.5mm} 
+    $\alpha_1=\{userId:\mathtt{"Lori"},\ action:\mathtt{login},\ time:\mathtt{17.0}\}$ \\$\alpha_2=\{userId:\mathtt{"Lori"},\ action:\mathtt{logout},\ time:\mathtt{150.0}\}$\\ 
+    \vspace{2.5mm}
+    $satisfies$ $M_2$.
+    \end{ex}
+\end{frame}
+
+\begin{frame}[allowframebreaks]{Syntax}
+    Each specification $S$ is made up of an observer part $O$ and a declaration part $D$.
     \begin{align*}
-        & \underline{\text{min}}\ \mathtt{EvLogout}(\underline{\text{string}}\ k) = (action = \mathtt{logout}\land userId = k) \lor \bigcirc \mathtt{EvLogout}(k) \\
-        & \underline{\text{mon}}\ M_2 = ((action = \mathtt{login})\rightarrow \mathtt{EvLogout}(userId)) 
+    S &::= D \, O \\
+    D &::= R^* \\
+    O &::= M^* \\
+    R &::= \{\underline{\text{max}}\ |\ \underline{\text{min}} \}\ N(T_1\ x_1, \ldots, T_n\ x_n) = F \\
+    M &::= \underline{\text{mon}}\ N = F \\
+    T &::= \underline{\text{Form}}\ |\ \text{primitive type} \\
+    F &::= \text{expression}\ |\ \underline{\text{true}}\ |\ \underline{\text{false}}\ |\ \neg F\ |\ F_1 \land F_2\ |\ F_1 \lor F_2\ |\\
+            &F_1 \rightarrow F_2\ |\ \bigcirc F\ |\ \odot F\ |\ F_1 \cdot F_2\ |\ N(F_1, \ldots, F_n)\ |\ x_i
     \end{align*}
-    Once the rules are created, the monitor will evaluate and update the monitored formula $M_2$. 
-    \vspace{2.5mm}
+    Let us focus on some details.\\
+    Definitions can start with different keywords:
+    \begin{itemize}
+        \item $\underline{\text{mon}}$: specifies the EAGLE formulas to be monitored and cannot have a recursive definition; as already told, these kind of rules will evolve as new events appear.
+        \item $\underline{\text{max}}$: defines \textit{safety properties} (nothing bad ever happens) and have a maximal interpretation.
+        \item $\underline{\text{min}}$: defines \textit{liveness properties} (something good eventually happens) and have a minimal interpretation.
+    \end{itemize}
 
-    A possibile trace $\sigma=\alpha_1,\alpha_2$, where $\alpha_1=\{userId:\mathtt{"Lori"},\ action:\mathtt{login},\ time:\mathtt{17.0}\}$ and $\alpha_2=\{userId:\mathtt{"Lori"},\ action:\mathtt{logout},\ time:\mathtt{150.0}\}$ can satisfy $M_2$.
 \end{frame}

### 2023-09-11 Presentation/presentation.tex
update

@@
 
+    \begin{block}{Interpretation note.}    
+        The difference between maximal and minimal becomes important only when we are evaluating the at the boundaries of a trace; $\underline{\text{max}}$ rules evaluates at $\underline{\text{true}}$ at initial and final istants, while $\underline{\text{min}}$ rules evaluates at $\underline{\text{false}}$.
+    \end{block}
+
 \end{frame}

### 2023-09-11 Presentation/presentation.tex
update

--- 2023-09-11 Presentation/presentation.tex
@@
 
+    \begin{block}{Note.}    
+        The difference between maximal and minimal interpretation becomes important only when we are evaluating the at the boundaries of a trace; $\underline{\text{max}}$ rules evaluates at $\underline{\text{true}}$ at initial and final istants, while $\underline{\text{min}}$ rules evaluates at $\underline{\text{false}}$.
+    \end{block}
+
 \end{frame}

### 2023-09-11 Paper/main.tex
update

@@
 
-\textbf{Semantics.} The semantics of the logic is defined in terms of the \textit{satisfaction} relation $\models$ which defines whether a finite execution trace $\sigma$ satisfy the specification $\varphi=D \, O$.
+\textbf{Semantics.} The semantics of the logic is defined in terms of the \textit{satisfaction} relation $\models$ which defines whether a finite execution trace $\sigma$ satisfies the specification $\varphi=D \, O$.
 The $i$'th state $s_i$ of a trace $\sigma$ is denoted by $\sigma(i)$.

### 2023-09-11 Presentation/presentation.tex
update

@@
 This property turns out to be crucial for succinctly expressing executions of attack-safe systems.
-
     \vspace{2.5mm}
     EAGLE operates with \textit{finite trace} semantics, meaning it checks formula satisfaction only at the end of a trace.
-However, in intrusion detection where event sequences can be infinite, the goal is to trigger an alarm as soon as a property is violated, thus MONID continuously checks the formula's satisfaction status after each event.
+However, in intrusion detection where event sequences can be infinite, the goal is to trigger an alarm as soon as a property is violated, thus MONID continuously checks the formula's satisfaction status after each event.\\
+    \vspace{2.5mm}
 
-    Let us start with an example; we want to express the property "\textit{Whenever there is a login by any user x, then eventually the user x logs out}".
-In EAGLE we can do it with the following rules:
+    Let us start with an example.
     \begin{ex} 
+    We want to express the property "\textit{Whenever there is a login by any user x, then eventually the user x logs out}". In EAGLE we can do it with the following rules:\\
     \vspace{2.5mm}
-    $ \underline{\text{min}}\ \mathtt{EvLogout}(\underline{\text{string}}\ k) = (action = \mathtt{logout}\land userId = k) \lor \bigcirc \mathtt{EvLogout}(k) $\\
-    $  \underline{\text{mon}}\ M_2 = ((action = \mathtt{login})\rightarrow \mathtt{EvLogout}(userId)) $\\
+    $\quad \underline{\text{min}}\ \mathtt{EvLogout}(\underline{\text{string}}\ k) = (action = \mathtt{logout}\land userId = k) \lor \bigcirc \mathtt{EvLogout}(k) $\\
+    $\quad \underline{\text{mon}}\ M_2 = ((action = \mathtt{login})\rightarrow \mathtt{EvLogout}(userId)) $\\
     \vspace{2.5mm}
@@
     \vspace{2.5mm}
-    $satisfies$ $M_2$.
+    satisfies $M_2$.
     \end{ex}
@@
     \end{block}    
+\end{frame}
+
+\begin{frame}[allowframebreaks]{Semantics}
+    The semantics of the logic is defined in terms of the \textit{satisfaction} relation $\models$ which defines whether a finite execution trace $\sigma$ satisfies the specification $\varphi=D \, O$.\\
+    \vspace*{5mm}
+    Given a trace $\sigma$ and a specification $D \, O$, satisfaction is defined as follows:
+    \[
+    \sigma \models D \, O \text{ iff } \forall (\underline{\text{mon}}\ N = F) \in O.\ \sigma ,1 \models_D \, F
+    \]
+    In other words, a trace satisfies a specification if it fulfills each monitored formula ($\underline{\text{mon}}$) when monitoring from position 1, which is the initial state.\\
+    \vspace*{5mm}
+    In the following slide we will present the definition of the satisfaction relation $\models_D \subseteq (Trace \times \mathbb{N}) \times \underline{\text{Form}}$ for a set of rule definitions $D$.
+
+    \includegraphics[scale=0.22]{images/semantics.png}
+\end{frame}
+
+\begin{frame}[allowframebreaks]{Evaluation algorithm}
+    As we can see, a function called $evaluate$ emerges; there are additional functions as well. Let us list them.
+    \begin{itemize}
+        \item $evaluate: \underline{\text{Form}} \times \underline{\text{State}}$. This function is used to evolve at every event the monitored formulas.
+        \item $update: \underline{\text{Form}} \times \underline{\text{State}} \to \underline{\text{Form}}$.
+This function is used by $evaluate$ to pre-evaluate a formula if it is guarded by the previous operator $\odot$
+        \item $value : \underline{\text{Form}} \rightarrow \{ \underline{\text{true}}, \underline{\text{false}} \}$.
+This function returns true if and only if $\sigma,|\sigma|+1 \models F$ (or $\sigma,0 \models F$), and returns false otherwise.
+    \end{itemize}
+    \begin{block}{Note.}
+        Formally, a formula $F' = evaluate(F, \alpha_i)$ is created when a formula $F$ is evaluated at an event $\alpha_i = \sigma(i)$ and has the condition that $\sigma, i \models F$ if and only if $\sigma, i+1 \models F'$.
+With $F$ being the evolved formula, we compute the boolean function $value(F)$ at the conclusion of the trace.
+    \end{block}
+
+    \newpage
+
+    Better details are in the paper; let us introduce an example.
+    \begin{ex}
+        Given the following rules\\
+        \vspace{2.5mm}
+        $\quad\underline{\text{max}}\ \mathtt{Always}(\underline{\text{Form}}\ F)  = F \land \bigcirc \mathtt{Always}(F)$ \\
+        $\quad\underline{\text{min}}\ \mathtt{EvTimedLogout}(\underline{\text{string }} k, \underline{\text{double }} t, \underline{\text{double }} \delta) = (\text{time} - t \leq \delta)$ \\
+        $\quad\quad\quad\quad\quad \land ((action = \mathtt{logout} \land userId = k) \lor \bigcirc \mathtt{EvTimedLogout}(k, t, \delta))$ \\
+        $\quad\underline{\text{mon}}\ M_3  = \mathtt{Always}((action = \mathtt{login}) \rightarrow \mathtt{EvTimedLogout}(userId,time, 100))$ \\
+        \vspace{2.5mm}
+        and a trace $\sigma=\alpha_1,\alpha_2$, where $\alpha_1=\{userId:\mathtt{"Lori"},\ action:\mathtt{login},\ time:\mathtt{17.0}\}$ and $\alpha_2=\{userId:\mathtt{"Lori"},\ action:\mathtt{logout},\ time:\mathtt{150.0}\}$.\\ $M_3$ gets modified, at first step, in \\
+        \vspace{2.5mm}
+        $\quad evaluate(M_3,\alpha_1)  = \mathtt{EvTimedLogout("Lori",17.0, 100)}$ \\ 
+        $\quad\quad\quad\quad\quad \land\ \mathtt{Always}((action = \mathtt{login}) \rightarrow \mathtt{EvTimedLogout}(userId,time, 100))$.
+
+    \end{ex}
 

### 2023-09-11 Presentation/presentation.tex
update

@@
     \newpage
-
-    Better details are in the paper; let us introduce an example.
+    More details are in the paper; let us introduce an example.
     \begin{ex}
         Given the following rules\\
-        \vspace{2.5mm}
+        \vspace{1.5mm}
         $\quad\underline{\text{max}}\ \mathtt{Always}(\underline{\text{Form}}\ F)  = F \land \bigcirc \mathtt{Always}(F)$ \\
@@
         $\quad\underline{\text{mon}}\ M_3  = \mathtt{Always}((action = \mathtt{login}) \rightarrow \mathtt{EvTimedLogout}(userId,time, 100))$ \\
-        \vspace{2.5mm}
-        and a trace $\sigma=\alpha_1,\alpha_2$, where $\alpha_1=\{userId:\mathtt{"Lori"},\ action:\mathtt{login},\ time:\mathtt{17.0}\}$ and $\alpha_2=\{userId:\mathtt{"Lori"},\ action:\mathtt{logout},\ time:\mathtt{150.0}\}$.\\ $M_3$ gets modified, at first step, in \\
-        \vspace{2.5mm}
+        \vspace{1.5mm}
+        and a trace $\sigma=\alpha_1,\alpha_2$ , where\\\vspace{1.5mm} $\quad\alpha_1=\{userId:\mathtt{"Lori"},\ action:\mathtt{login},\ time:\mathtt{17.0}\}$\\ $\quad\alpha_2=\{userId:\mathtt{"Lori"},\ action:\mathtt{logout},\ time:\mathtt{150.0}\}$.\\ \vspace{1.5mm}$M_3$ gets modified, at first step, in \\
+        \vspace{1.5mm}
         $\quad evaluate(M_3,\alpha_1)  = \mathtt{EvTimedLogout("Lori",17.0, 100)}$ \\ 
         $\quad\quad\quad\quad\quad \land\ \mathtt{Always}((action = \mathtt{login}) \rightarrow \mathtt{EvTimedLogout}(userId,time, 100))$.
-
-    \end{ex}
-
+    \end{ex}    %continuazione?
 \end{frame}

### 2023-09-11 Paper/main.tex
update

@@
 \end{align*}
-By combining the definitions for the future and past time LTLs defined above, we obtain a temporal logic over the future, present and past, in which one can freely intermix the future and past time modalities; thus, we can state that EAGLE is expressive at least as LTL+P.
+By combining the definitions for the future and past time LTL as defined above, we obtain a temporal logic over the future, present and past, in which one can freely intermix the future and past time modalities; thus, we can state that EAGLE is expressive at least as LTL+P.
 % a differenza di ltl non salva tutta la sequenza ma solo una formula che riassume tutto. La seconda componente non è uno stato (insieme di cose vere) ma un record

### 2023-09-11 Presentation/presentation.tex
update

@@
 
+\begin{frame}{Relationship with LTL+P}
+    We can define both \textbf{Future Time LTL} and \textbf{Past Time LTL} operators. As example, let us show the future ones.
+    \begin{align*}
+        & \underline{\text{min}}\ \mathtt{Next}(\underline{\text{Form}}\ F) = \bigcirc F \\
+        & \underline{\text{max}}\ \mathtt{Always}(\underline{\text{Form}}\ F) = F \land \bigcirc \mathtt{Always}(F) \\
+        & \underline{\text{min}}\ \mathtt{Eventually}(\underline{\text{Form}}\ F) = F \lor \bigcirc \mathtt{Eventually}(F) \\
+        & \underline{\text{min}}\ \mathtt{Until}(\underline{\text{Form}}\ F_1, \underline{\text{Form}}\ F_2) = F_2 \lor (F_1 \land \bigcirc \mathtt{Until}(F_1,F_2))
+    \end{align*}
+    By combining the definitions for the future and past time LTL as defined above, we obtain a temporal logic over the future, present and past, in which one can freely intermix the future and past time modalities; thus, we can state that EAGLE is expressive at least as LTL+P.
+
+\end{frame}
+
 \section{Conclusions}

### 2023-09-11 Presentation/presentation.tex
update

@@
     \end{align*}
-    By combining the definitions for the future and past time LTL as defined above, we obtain a temporal logic over the future, present and past, in which one can freely intermix the future and past time modalities; thus, we can state that EAGLE is expressive at least as LTL+P.
+    By combining the definitions for the future and past time LTL as defined above, we obtain a temporal logic over the future, present and past, in which one can freely intermix the future and past time modalities;\\thus, we can state that EAGLE is expressive at least as LTL+P.
 

### 2023-09-12 Paper/main.tex
update

@@
 \end{align*}
-as we can see, the \textit{type} field of a record schema describes the type of network packet, while the \textit{ip} field is the destination of the packet.The $\mathtt{SmurfAttack}$ rule determines whether the packet is a ping packet and whether the source's IP address is a broadcast ip, i.e.
+as we can see, the \textit{type} field of a record schema describes the type of network packet, while the \textit{ip} field is the destination of the packet.
+The $\mathtt{SmurfAttack}$ rule determines whether the packet is a ping packet and whether the source's IP address is a broadcast ip, i.e.
 all the host bits are set to 1.

### 2023-09-12 Presentation/presentation.tex
update

@@
 
+\section{Attacks detection}
+\setLayout{mainpoint}
+\setBGColor{LightOrange}
+\begin{frame}{}
+    \frametitle{Attacks detection}
+\end{frame}
+\setLayout{vertical}
+
+\begin{frame}
+    \frametitle{Smurf Attacks}
+    A \textit{smurf attack} is a DDoS attack targeting computer networks by sending large volumes of ICMP echo request packets to a broadcast address.
+The attacker spoofs the source IP address to make it appear as originating from the victim's IP address, causing all devices to respond with ICMP echo replies.\\
+    \vspace{2.5mm}
+    In this example we will use \textit{tcpdump} to obtain the logs from the network interface; the formula that describes the absence of the attack is the following.
+    \begin{align*}
+        & \underline{\text{max}}\ \mathtt{SmurfAttack()} = (type = \text{"ICMP"})\land isBroadcast(ip) \\
+        & \underline{\text{mon}}\ \mathtt{SmurfSafety} =\ \mathtt{Always}(\neg\ \mathtt{SmurfAttack()}) 
+    \end{align*}
+
+\end{frame}
+
 \section{Conclusions}

### 2023-09-12 Presentation/presentation.tex
update

@@
     \vspace{2.5mm}
-    In this example we will use \textit{tcpdump} to obtain the logs from the network interface; the formula that describes the absence of the attack is the following.
+    In this example we will use \textit{tcpdump} to obtain the logs from the network interface; the formula that describes the absence of the attack is the following:
     \begin{align*}
@@
     \end{align*}
+    where \textit{isBroadcast}() checks whether the source’s IP address is a broadcast ip, i.e. all the host bits are set to 1.
+\end{frame}
 
+\begin{frame}
+    \frametitle{Cookie-stealing Attacks}
+    A \textit{cookie} is a technique used by web application servers to track client-specific session data; these tools are automatically included in client requests.
+When a malicious user uses an old cookie provided to a different IP address to take over a session, it is considered an attack.\\ \vspace{2.5mm} At this point we need to look at web-server log; the formula that describes the absence of the attack is the following:
+    \begin{align*}
+        & \underline{\text{min}}\ \mathtt{Hijack}(\underline{\text{string}}\ c,\underline{\text{string}}\ i) = ((name=c)\land\neg(ip=i))\lor\odot\ \mathtt{Hijack}(c,i) \\
+        & \underline{\text{mon}}\ \mathtt{CookieSafe} =\ \mathtt{Always}(\neg\mathtt{Hijack}(name,ip)) 
+    \end{align*}
+    where a single cookie is identified from $c$ and it can be associated only at a single $ip$.
+\end{frame}
+
+\begin{frame}
+    \frametitle{Multi-domain Buffer Overflows attacks}
+    A \textit{buffer-overflow} attack uses memory manipulation to overflow a buffer, modifying an address to point to malicious code.
+Data from network and web server access logs are analyzed to determine when the attack occurred and if no matching log record is found within a timeout, the attack is successful.
+To capture this scenario we use the following formulas:
+    \begin{align*}
+        & \underline{\text{min}}\ \mathtt{EventuallyClosed}(\underline{\text{long}}\ t,\underline{\text{long}}\ d,\underline{\text{string}}\ i1,\underline{\text{string}}\ i2) = (time-t<d)\\
+        & \quad\quad\land ((ip1=i1\land ip2=i2\land log=web\land type=closed)\\
+        & \quad\quad\quad\lor \bigcirc\mathtt{EventuallyClosed}(t,d,i1,i2) ) \\
+        & \underline{\text{mon}}\ \mathtt{BufferSafe} =\ \mathtt{Always}((log=network\land type=binary) \\
+        & \quad\quad\rightarrow\mathtt{EventuallyClosed}(time,100,ip1,ip2))
+    \end{align*}
+    where a session, identified by $(ip1,ip2)$, must end within $100$ seconds if there is a $binary$ access.
+\end{frame}
+
+\begin{frame}
+    \frametitle{Password guessing attacks}
+    If the system accepts telnet connections for remote login, an attacker can easily compromise a user's password by guessing multiple passwords for a specific user name.
+We can detect with the following:
+    \begin{align*}
+        & \underline{\text{max}}\ \mathtt{Failure}() = (type=login)\land \neg success\\
+        & \underline{\text{max}}\ \mathtt{Guess}(\underline{\text{string}}\ i) = (ip=i)\land \mathtt{Failure}()\\
+        & \underline{\text{max}}\ \mathtt{Counter}(\underline{\text{long}}\ t, \underline{\text{long}}\ d, \underline{\text{int}}\ c,\underline{\text{string}}\ i,\underline{\text{int}}\ C) = (time-t<d)\\
+        & \quad\quad\rightarrow((\mathtt{Guess}(i)\rightarrow(c\leq C\land \bigcirc\mathtt{Counter}(t,d,c+1,i,C)))  \\
+        & \quad\quad\land(\neg \mathtt{Guess}(i)\rightarrow\mathtt{Counter}(t,d,c,i,C))) \\
+        & \underline{\text{mon}}\ \mathtt{PassGuessSafe} =\ \mathtt{Always}(\mathtt{Failure}()\rightarrow\mathtt{Counter}(time,300,1,ip,3)) 
+    \end{align*}
+    where an attacks is detected if a password has been incorrectly guessed for $3$ times within $300$ seconds.
 \end{frame}

### 2023-09-12 Presentation/presentation.tex
update

@@
     \frametitle{Multi-domain Buffer Overflows attacks}
-    A \textit{buffer-overflow} attack uses memory manipulation to overflow a buffer, modifying an address to point to malicious code.
+    A \textit{buffer overflow} attack uses memory manipulation to overflow a buffer, modifying an address to point to malicious code.
 Data from network and web server access logs are analyzed to determine when the attack occurred and if no matching log record is found within a timeout, the attack is successful.
@@
     where an attacks is detected if a password has been incorrectly guessed for $3$ times within $300$ seconds.
+\end{frame}
+
+\begin{frame}
+    \frametitle{Port Scanning attacks}
+    \textit{Port scanning} is a method used to identify exploitable communication channels in networks; it involves probing network ports and keeping useful information for attacks.
+The victim machine suspect port scans as malicious when the counter of reached ports exceeds a specific threshold.
+The formulas are the following:
+    \begin{align*}
+        & \underline{\text{max}}\ \mathtt{NewPort}(\underline{\text{string}}\ i1,\underline{\text{string}}\ i2,\underline{\text{Set}}\ S) = (i1=ip1)\land(i2=ip2)\land (port \notin S)\\
+        & \underline{\text{max}}\ \mathtt{Counter}(\underline{\text{long}}\ t, \underline{\text{long}}\ d, \underline{\text{int}}\ c,\underline{\text{string}}\ i1,\underline{\text{string}}\ i2,\underline{\text{Set}}\ S,\underline{\text{int}}\ C) = (time-t<d)\\
+        & \quad\quad\rightarrow((\mathtt{NewPort}(i1,i2,S)\rightarrow(c\leq C\land \bigcirc\mathtt{Counter}(t,d,c+1,i1,i2,S\cup\{port\},C)))  \\
+        & \quad\quad\land(\neg \mathtt{NewPort}(i1,i2,S)\rightarrow\mathtt{Counter}(t,d,c,i1,i2,S,C))) \\
+        & \underline{\text{mon}}\ \mathtt{PortScanSafe} =\ \mathtt{Always}(\mathtt{Counter}(time,100,1,ip1,ip2,\{port\},10)) 
+    \end{align*}
+    where we checks if the number of port scans between $ip1,ip2$ does not exceed a threshold of $10$ within $100$ seconds.
 \end{frame}

### 2023-09-12 Presentation/presentation.tex
update

@@
     \frametitle{Conclusions}
-
+    In conclusion:
+    \begin{itemize}
+        \item We have introduced the MONID framework and explained how it works.
+        \item We then proceeded introducing a new temporal logic named EAGLE, which serves as the foundation for the MONID framework.
+        \item We have presented some example of formulas which aims at recognize some attacks patterns.
+    \end{itemize}
+    \vspace{2.5mm}
+    Due to the little overhead involved in the computation and the great performances of MONID when utilized offline, we can state that this method should still be useful in theory today.
 \end{frame}

### 2023-09-12 Paper/main.tex
update

@@
 The experiment was done using offline mode, against some of the attacks presented in Section~\ref{attacks}.
+More specifically, against the smurf, port-sweep and password-guessing attacks; the experiments detected 5 password-guess attack and 2 port-sweep attack in the logs and the performance overheads for monitoring the Port-Sweep and Password-Attacks are reported in Figure~\ref{graph1} and~\ref{graph2}, respectively.
+The X-axis in both graphs shows the number of events we are monitoring, as they are obtained from our logs.
+Each data point is the average overhead calculated for intervals of 10 and 1000 events respectively.
+The Y-axis plots the ratio between the time spent by the monitor vs.
+the time between the generation of the events in the actual log.
+As long as this ratio is less than 1, our monitoring is feasible.
+The results, show that the average overhead, is around $0.009$ for Port-Sweep attacks and $0.016$ for Password-Guessing attacks.
 Although the used hardware is now outdated and obsolete, the overhead was rather modest, thus this method should still be theoretically helpful today.
+
+\begin{figure}[!htb]
+    \begin{center}
+        \begin{minipage}{0.45\textwidth}
+            \centering
+            \includegraphics[width=\linewidth]{images/graph 1.png}
+            \caption{Performances overhead of Port-Sweep attacks.}\label{graph1}
+        \end{minipage}
+        \hfill
+        \begin{minipage}{0.45\textwidth}
+            \centering
+            \includegraphics[width=\linewidth]{images/graph 2.png}
+            \caption{Performances overhead of Password-Guessing attacks.}\label{graph2}
+        \end{minipage}
+    \end{center}
+\end{figure}
 

### 2023-09-12 Presentation/presentation.tex
update

@@
 
+\begin{frame}
+    \frametitle{Performances}
+    MONID was evaluated against smurf, port-sweep, and password-guessing attacks using DARPA Intrusion Detection Evaluation dataset in offline mode, detecting 5 password-guess attacks and 2 port-sweep attacks.
+    \begin{figure}[!htb]
+        \begin{center}
+            \begin{minipage}{0.45\textwidth}
+                \centering
+                \includegraphics[width=\linewidth]{images/graph 1.png}
+                \caption{Performances overhead of Port-Sweep attacks.}\label{graph1}
+            \end{minipage}
+            \hfill
+            \begin{minipage}{0.45\textwidth}
+                \centering
+                \includegraphics[width=\linewidth]{images/graph 2.png}
+                \caption{Performances overhead of Password-Guessing attacks.}\label{graph2}
+            \end{minipage}
+        \end{center}
+    \end{figure}
+
+\end{frame}
+
 \section{Conclusions}

### 2023-09-16 Presentation/presentation.tex
update

@@
     As a result, rules in EAGLE give us the power to create specific temporal operators as well as to bind and modify data.
-This property turns out to be crucial for succinctly expressing executions of attack-safe systems.
+This property turns out to be crucial for succinctly expressing executions of attack-safe systems.\\
     \vspace{2.5mm}
@@
     Due to the little overhead involved in the computation and the great performances of MONID when utilized offline, we can state that this method should still be useful in theory today.
+
+    \vspace{2.5mm}
+    Attacks nowadays become more and more complex but once a pattern is found, we can represent it with EAGLE.
 \end{frame}

### 2023-09-16 Paper/paper.tex
update

--- 2023-09-12 Paper/main.tex
@@
 
-There are three Intrusion Detection System approaches, according to the literature: \textit{signature-based}\cite{gao2014cyber}, \textit{anomaly-based}\cite{jyothsna2011review} and \textit{hybrid}\cite{garg2016hybrid}.
-The first approach aims to identify patterns and match them with known signs of intrusions relying on a database of previous intrusions.
-If activity within the network matches the “signature” of an attack or breach from the database, the detection system creates an alert.
+According to the literature, there are three approaches to intrusion detection systems: \textit{signature-based}\cite{gao2014cyber}, \textit{anomaly-based}\cite{jyothsna2011review} and \textit{hybrid}\cite{garg2016hybrid}.
+The first approach aims to identify patterns and match them with known signs of intrusions, relying on a database of previous intrusions.
+If activity within the network matches the “signature” of an attack or breach in the database, the detection system creates an alert.
 This approach has a low false-alarm rate, but it requires us to know the patterns of security attacks in advance and previously unknown attacks would go undetected.
-In contrast, \textit{anomaly-based} is capable to detect new attacks since an alarm is raised if an observed behavior deviates significantly from pre-learned normal behavior.
-Finally, a \textit{hybrid system} combines the best of both worlds by looking at patterns and one-off events, a Hybrid Intrusion Detection system can flag new and existing intrusion strategies.
+In contrast, \textit{anomaly-based} is capable of detecting new attacks since an alarm is raised if an observed behavior deviates significantly from pre-learned normal behavior.
+Finally, a \textit{hybrid system} combines the best of both worlds by looking at patterns and one-off events.
+A Hybrid Intrusion Detection system can flag new and existing intrusion strategies.
 
@@
 An intrusion alarm is triggered if, for any reason, the updated formula turns out to be false; as a result, the entire process operates in real-time.
-We'll see how the implementation, MONID, examines the event stream to see if the monitored formula (which is used to hold the pertinent summary of the system) is being broken.
+We'll see how the implementation, MONID, examines the event stream to see if the monitored formula (which is used to hold the pertinent summary of the system) is broken.
 
@@
 %%%%%   DA RISCRIVERE ALLA FINE
-In Section 2, we'll deeply explore EAGLE, covering everything from its syntax to semantics and the evaluation algorithm, with the add of practical examples.
+In Section 2, we'll deeply explore EAGLE, covering everything from its syntax to semantics and the evaluation algorithm, with the addition of practical examples.
 Moving on to Section 3, we'll delve into the MONID framework, which employs EAGLE to construct an intrusion detection system.
@@
 \subsection{Basics}\label{sec:ltl}
-In EAGLE recursion definitions are supported; as exmaple, in the current framework we can build the following definitions:
+In EAGLE recursion definitions are supported; for example, in the current framework we can build the following definitions:
 \begin{align*}
@@
 As we can see, rules are parameterized by an EAGLE formula (of type $\underline{\text{Form}}$), which means that we will be able to write EAGLE formulas such as $\mathtt{Always}(\mathtt{Eventually}(x>0))$.
-Also, the $\mathtt{Always}$ operator is defined as maximal solution of the equation $X=F\land \bigcirc X$, while the $\mathtt{Eventually}$ operator represents the minimal solution to the equation $X=F\lor \bigcirc X$.
+Also, the $\mathtt{Always}$ operator is defined as the maximal solution of the equation $X=F\land \bigcirc X$, while the $\mathtt{Eventually}$ operator represents the minimal solution to the equation $X=F\lor \bigcirc X$.
 
@@
     An example of event could be: $\{userId:\mathtt{"Lori"},\ action:\mathtt{login},\ time:\mathtt{20}\}$
-    \item For each attack, there is a formula $F$ which specifies the absence of it.
+    \item For each attack, there is a formula $F$ that specifies the absence of it.
 \end{enumerate}
@@
 \textbf{Syntax.}
-A specification $S$ is made up of an observer part $O$ and a declaration part $D$.
+A specification $S$ is made up of an observer's part $O$ and a declaration's part $D$.
 $O$ is made up of zero or more monitor definitions $M$, which define what will be watched, while $D$ is made up of zero or more rule definitions $R$.
@@
 or formulas of type \underline{Form}.
-Definitions starting with keyword $\underline{\text{mon}}$ specifies the EAGLE formulas to be monitored and cannot have a recursive definition.
-We will see that these kind of rules will evolve as new events appear.
-A term indicating whether the interpretation is maximal ($\underline{\text{max}}$) or minimal ($\underline{\text{min}}$) comes before a rule specification $R$.
+Definitions starting with the keyword $\underline{\text{mon}}$ specify the EAGLE formulas to be monitored and cannot have a recursive definition.
+We will see that these kinds of rules evolve as new events occur.
+A term indicating whether the interpretation is maximal ($\underline{\text{max}}$) or minimal ($\underline{\text{min}}$) comes before the rule specification $R$.
 While minimal rules establish \textit{liveness properties} (something good eventually happens), maximal rules define \textit{safety properties} (nothing bad ever happens); we shall see that the difference becomes important only when we are evaluating the at the boundaries of a trace.
@@
 if and only if the rule is true at some given event $\alpha_i$. 
-Once the sequence has been completely analyzed, we will obtain a big disjunction (in case of a \underline{\text{min}}) or a big conjunction (in case of a \underline{\text{max}}) and the final value will close the evalutation.
+Once the sequence has been completely analyzed, we will obtain a big disjunction (in the case of a \underline{\text{min}}) or a big conjunction (in the case of a \underline{\text{max}}) and the final value will close the evaluation.
 Continuing the example, the extended formula will be:
@@
 \end{align*}
-and the bold $\mathbf{Property(F)}$ will be evaluated to false, since it is the last of a big disjunction.
+and the bold $\mathbf{Property(F)}$ will be evaluated as false, since it is the last of a big disjunction.
 \subsection{Evaluation algorithm}\label{algo}
@@
             \centering
-            \includegraphics[width=\linewidth]{images/graph 1.png}
+            \includegraphics[width=1.1\linewidth]{images/graph 1.png}
             \caption{Performances overhead of Port-Sweep attacks.}\label{graph1}
@@
             \centering
-            \includegraphics[width=\linewidth]{images/graph 2.png}
+            \includegraphics[width=1.1\linewidth]{images/graph 2.png}
             \caption{Performances overhead of Password-Guessing attacks.}\label{graph2}
@@
 According to the formula below, a specific cookie must always be utilized by the same IP address.
-
 \begin{align*}

### 2023-09-16 Paper/paper.tex
update

@@
 \subsection{Evaluation algorithm}\label{algo}
-As already mentioned, monitored formulas ($\underline{\text{mon}}$) evolve at every event, to store relevant informations about past istants.
+As already mentioned, monitored formulas ($\underline{\text{mon}}$) evolve at every event to store relevant information about past events.
 The evolved formula's value is determined at the conclusion of the event series; if it is true, the formula is satisfied by the event sequence; otherwise, it is violated.
@@
 With $F$ being the evolved formula, we compute the boolean function $value(F)$ at the conclusion of the trace.
-The definition of the function $evaluate: \underline{\text{Form}} \times \underline{\text{State}} \to \underline{\text{Form}}$ uses an auxiliary function $update: \underline{\text{Form}} \times \underline{\text{State}} \to \underline{\text{Form}}$ to pre-evaluate a formula if it is guarded by the previous operator $\bigodot$.
+The function $evaluate: \underline{\text{Form}} \times \underline{\text{State}} \to \underline{\text{Form}}$ uses an auxiliary function $update: \underline{\text{Form}} \times \underline{\text{State}} \to \underline{\text{Form}}$ to pre-evaluate a formula if it is guarded by the previous operator $\bigodot$.
 Note that $\underline{\text{Form}}$ in the practical use represents the monitored formula, while $\underline{\text{State}}$ represents the current event.
@@
 
-\textbf{Calculus.} According to~\cite{barringer2004rule}, the $evaluate$, $update$ and $value$ functions are defined a priori for all operators except for the
-rule application, for which get generated based on the definition of rules in the specification.
+\textbf{Calculus.} According to~\cite{barringer2004rule}, the $evaluate$, $update$ and $value$ functions are defined a priori for all operators except for the rule application, which gets generated based on the definition of rules in the specification.
 The definitions of these functions on the different primitive operators (except the previous operator $\odot$) are the following ($op\text{ can be }\lor , \land, \rightarrow $):
@@
 \end{align*}
-At the second event the predicate $(time-t \leq \delta)$ gets instantiated to $(150.0-17.0 \leq 100)$ which is false; thus the whole formula becomes false, indicating that the two-event trace violates the property $M_3$.
+At the second event the predicate $(time-t \leq \delta)$ gets instantiated to $(150.0-17.0 \leq 100)$ which is false; thus, the whole formula becomes false, indicating that the two-event trace violates the property $M_3$.
 
@@
 System-level event data is delivered to MONID server either directly from relevant log files (offline) or online via application code that has been properly instrumented.
-In the online mode, MONID runs as a server that receives streams of events
-from various sources and to generate these events, the different logging modules are instrumented so that they filter and send relevant events to the server.
+In the online mode, MONID runs as a server that receives streams of events from various sources and to generate these events, the different logging modules are instrumented so that they filter and send relevant events to the server.
 To create a single event trace, the server combines events from numerous sources according to timestamps and preprocesses them into an abstract intermediate form.

### 2023-09-16 Paper/paper.tex
update

@@
 \section{MONID: Structure analysis}
-MONID is a prototype which can detect intrusions either online or offline.
+MONID is a prototype that can detect intrusions either online or offline.
 System-level event data is delivered to MONID server either directly from relevant log files (offline) or online via application code that has been properly instrumented.
@@
 To create a single event trace, the server combines events from numerous sources according to timestamps and preprocesses them into an abstract intermediate form.
-Note that sometimes it is necessary to watch events from multiple sources in order to detect certain attacks; then, MONID monitor checks this event trail against a predetermined criteria, and if the specification is broken, an intrusion alarm is raised.
-While, in the offline mode, MONID reads various log files and sends an event corresponding each log entry to the server.
-The server then processes the event stream as before to detect intrusion.
+Note that sometimes it is necessary to watch events from multiple sources in order to detect certain attacks; then, the MONID monitor checks this event trail against a predetermined criteria, and if the specification is broken, an intrusion alarm is raised.
+While in offline mode, MONID reads various log files and sends an event corresponding to each log entry to the server.
+The server then processes the event stream as before to detect an intrusion.
 
@@
 
-Some evaluations were also have been done in~\cite{naldurg2004temporal}, using the DARPA Intrusion Detection Evaluation dataset~\cite{mchugh2000testing}.
-The experiment was done using offline mode, against some of the attacks presented in Section~\ref{attacks}.
-More specifically, against the smurf, port-sweep and password-guessing attacks; the experiments detected 5 password-guess attack and 2 port-sweep attack in the logs and the performance overheads for monitoring the Port-Sweep and Password-Attacks are reported in Figure~\ref{graph1} and~\ref{graph2}, respectively.
+Some evaluations were also done in~\cite{naldurg2004temporal}, using the DARPA Intrusion Detection Evaluation dataset~\cite{mchugh2000testing}.
+The experiment was done using offline mode against some of the attacks presented in Section~\ref{attacks}.
+More specifically, against the smurf, port-sweep and password-guessing attacks, the experiments detected 5 password-guess attack and 2 port-sweep attack in the logs and the performance overheads for monitoring the Port-Sweep and Password-Attacks are reported in Figure~\ref{graph1} and~\ref{graph2}, respectively.
 The X-axis in both graphs shows the number of events we are monitoring, as they are obtained from our logs.
@@
 As long as this ratio is less than 1, our monitoring is feasible.
-The results, show that the average overhead, is around $0.009$ for Port-Sweep attacks and $0.016$ for Password-Guessing attacks.
+The results show that the average overhead is around $0.009$ for Port-Sweep attacks and $0.016$ for Password-Guessing attacks.
 Although the used hardware is now outdated and obsolete, the overhead was rather modest, thus this method should still be theoretically helpful today.
@@
 \section{Example of attacks}\label{attacks}
-We give some examples of how EAGLE can be utilized to build formulas that correlate to desired execution trace characteristics of a system under observation in this section.
+In this section, we give some examples of how EAGLE can be utilized to build formulas that correlate to the desired execution trace characteristics of a system under observation.
 In this context, a trace $\sigma$ that violates this specification is referred to as an intrusion or an attack; to explain how the framework may be used, we use examples from actual attacks illustrated in~\cite{naldurg2004temporal}.
@@
 A smurf attack~\cite{zargar2009identification} is a type of Distributed Denial of Service (DDoS) attack that targets computer networks; in this scenario, the attacker sends a large volume of Internet Control Message Protocol (ICMP) echo request packets (commonly known as "ping" packets) to a network's broadcast address.
-This broadcast address is the address that reaches all devices on the targeted network.
+This broadcast address is the one that reaches all devices on the targeted network.
 The attacker spoofs the source IP address of these ICMP packets to make it appear as if they are originating from the victim's IP address and since the broadcast address sends the request to all devices on the network, all devices will respond with ICMP echo replies to the victim's IP address.
 This results in a flood of traffic overwhelming the victim's network and causing it to become unreachable or severely degraded in performance.
-To detect this kind of attacks, we need to work on network logs; we will use a network tool called \textit{tcpdump}.
+To detect this kind of attack, we need to work on network logs; we will use a network tool called \textit{tcpdump}.
 The monitored formula that describes the absence of the attack is the following:
@@
 These cookies that serve as authentication tokens are automatically included in requests by clients and in this scenario we assume that a session is exclusively identified by the client's IP address.
-When a malicious user uses an old cookie provided to a different IP address to take over a session, it is considered an attack.In order to monitor this attack we need to look at a web-server (application-level) log that contains a record of all sessions that the server participates in, along with session-specific state information.
-According to the formula below, a specific cookie must always be utilized by the same IP address.
+When a malicious user uses an old cookie provided by a different IP address to take over a session, it is considered an attack.
+In order to monitor this attack, we need to look at a web-server (application-level) log that contains a record of all sessions that the server participates in, along with session-specific state information.
+According to the formula below, a specific cookie must always be used by the same IP address.
 \begin{align*}
@@
 A trace that violates this formula therefore encodes a cookie-stealing attack.
-The rule $\mathtt{Hijack}$ checks wheter it occurs a situation in which we have the same coookie $c$ associated with two different ip $i$; if it's true then it means that a cookie has been stolen and reused by an attacker.
-With $\mathtt{CookieSafe}$ we want to check that it never occurs a situation just as the one described.
+The rule $\mathtt{Hijack}$ checks whether it occurs a situation in which we have the same cookie $c$ associated with two different ip $i$; if it's true, then it means that a cookie has been stolen and reused by an attacker.
+With $\mathtt{CookieSafe}$ we want to check that it never occurs a situation just like the one described.
 
@@
 In this case, we will integrate data from various log sources; to be more exact, we will examine both network and web server access logs to determine when a server buffer overflow attack has been performed.
-We search through network packets for binary data; then web server's access logs are examined to see if a corresponding event, in which the web server correctly terminates the connection after receiving some binary data, can be located.
+We search through network packets for binary data; then the web server's access logs are examined to see if a corresponding event, in which the web server correctly terminates the connection after receiving some binary data, can be located.
 If no matching log record is discovered within a predetermined timeout, it means that the buffer overflow attack was successful.
-The following formulas captures this scenario:
+The following formulas capture this scenario:
 \begin{align*} 
@@
 \end{align*}
-In each record we have the $log$ field which indicates the source of the log, the $time$ represents the istant in which the event occurs, $ip1$ represents the source address, $ip2$ represents the destination IP address and finally the $type$ field can be $binary, closed,\ldots$ We assume that a connection is identified by the tuple $(ip1,ip2)$.
+In each record we have the $log$ field which indicates the source of the log, the $time$ represents the instant in which the event occurs, $ip1$ represents the source address, $ip2$ represents the destination IP address and finally the $type$ field can be $binary, closed,\ldots$ We assume that a connection is identified by the tuple $(ip1,ip2)$.
 In order, the $\mathtt{EventuallyClosed}$ rule states that the connection between IP addresses $i1$ and $i2$ must end eventually within time $d$; finally, the monitor rule $\mathtt{BufferSafe}$ states that it is always the case that if there is an event of $binary$ access in the network log then eventually within time $100$ there must be a matching event in the webserver log that denotes the closing of the connection.
@@
 \textbf{Password guessing attacks.}
-Assuming the system accepts telnet connections to remote login; it won't take long until the password is compromised if an attacker is permitted to guess any amount of passwords for a specific user name.
+Assuming the system accepts telnet connections to remote login; it won't take long until the password is compromised if an attacker is permitted to guess any number of passwords for a specific user name.
 If a user guesses an incorrect password more than three times in a short amount of time, most systems end the telnet session.
@@
 The number of guesses is increased by one for each instance of an event signifying a failed login attempt from the same IP address within the timeout period $d$.
-The rule $\mathtt{Counter}$ also determines whether the local counter $c$ surpasses the threshold $C$ within time d, and if it does, the entire rule become false, indicating an attack.
-Finally, with the rule $\mathtt{PassGuessSafe}$ we assure that whenever there is a failed login from a specific IP address, we make sure that within time $300$ seconds the number of login-failures from the same IP address must be less than or equal to $3$.
+The rule $\mathtt{Counter}$ also determines whether the local counter $c$ surpasses the threshold $C$ within time d, and if it does, the entire rule becomes false, indicating an attack.
+Finally, with the rule $\mathtt{PassGuessSafe}$ we assure that whenever there is a failed login from a specific IP address, we make sure that within $300$ seconds the number of login-failures from the same IP address must be less than or equal to $3$.
 
@@
 According to~\cite{kanlayasiri2000rule}, Port scanning attack is a method for discovering exploitable communication channels that has been used for a long time.
-The key idea is to probe the network ports and then keep the information about them that are useful for an attack.
+The key idea is to probe the network ports and then keep the information about them that is useful for an attack.
 Hackers frequently employ a port scan approach to find gaps or weak spots in a network.
 Attackers can use a port scan attack to identify open ports and determine if they are accepting or rejecting data; additionally, it can show whether a company uses firewalls or other active security measures.
-Assuming that our server has multiple network inferfaces, thus multiple IP addresses; the victim machine can presume that port scans are malicious when the counter of reached ports overcomes a specific threshold in a short amount of time.
+Assuming that our server has multiple network interfaces, thus multiple IP addresses; the victim machine can presume that port scans are malicious when the counter of reached ports overcomes a specific threshold in a short amount of time.
 The formulas are the following:
@@
 in this case we have two IPs; $ip1$ is the source IP, while $ip2$ is the destination one.
-The parameterized $\mathtt{Counter}$ rule states that the number of port scans between a source and destination IP  address pair $(i1,i2)$ recorded do not surpass a threshold $C$ within time $d$.
-The rule $\mathtt{NewPort}$ checks if the port number,
-involved in any communication between the IP addresses $i1$ and $i2$, exists in the set $S$ of all port numbers (involved in all communications between $i1$ and $i2$) within the timeout period $d$.
+The parameterized $\mathtt{Counter}$ rule states that the number of port scans between a source and destination IP  address pair $(i1,i2)$ recorded does not surpass a threshold $C$ within time $d$.
+The rule $\mathtt{NewPort}$ checks if the port number involved in any communication between the IP addresses $i1$ and $i2$ exists in the set $S$ of all port numbers (involved in all communications between $i1$ and $i2$) within the timeout period $d$.
 
@@
 In conclusion, we have introduced a new temporal logic named EAGLE, which serves as the foundation for the MONID framework, which is a useful tool for system intrusion detection.
-MONID continuously monitor these attack specifications during system execution and when the observed execution contradicts the monitored formula, we detect an intrusion.
+MONID continuously monitors these attack specifications during system execution and when the observed execution contradicts the monitored formula, we detect an intrusion.
 We have showcased this by creating formulas to identify several common attacks and testing our monitoring algorithm on large event logs provided by DARPA.

### 2023-09-20 Presentation/presentation.tex
update

@@
     \end{block}    
+    \newpage
+    To better understand the mechanism, let us introduce an example.\\ Given the definition 
+    \begin{align*}
+        & \underline{\text{min}}\ \mathtt{Property}(\underline{\text{Form}}\ F) =F \lor \bigcirc \mathtt{Property}(F)
+    \end{align*} 
+    and the sequence $\sigma=\alpha_{1},\ldots ,\alpha_{n}$, $\mathtt{Property}$ will evaluate to \textit{true} 
+    \begin{align*}
+        & evaluate(\ldots evaluate(evaluate(\mathtt{Property},\alpha_1),\alpha_2),\ldots,\alpha_n)=true
+    \end{align*}
+    if and only if the rule is true at some given event $\alpha_i$. Once the sequence has been completely analyzed, we will obtain the following big disjunction: 
+    \begin{align*}
+        & F \lor \bigcirc(F \lor \bigcirc(\ldots F \lor \bigcirc \mathbf{Property(F)})) 
+    \end{align*}
+    $\mathbf{Property(F)}$ will be evaluated as false since it is the last of a big disjunction and we want to close it.
+
 \end{frame}
@@
     \end{align*}
-    where we checks if the number of port scans between $ip1,ip2$ does not exceed a threshold of $10$ within $100$ seconds.
+    where we checks if the number of (new) port scans between $ip1,ip2$ does not exceed a threshold of $10$ within $100$ seconds.
 \end{frame}

### 2023-09-20 Paper/paper.tex
update

@@
     \text{if } 1 \leq i \leq |\sigma| \text{ then:} \\
-    \quad \sigma, i \models_D F[ \overline{f} \mapsto \overline{F}, \overline{p} \mapsto\ evaluate(\overline{P}, \sigma(i))] \\
-    \quad \text{where } (N(\underline{\overline{\text{Form}}}\ \overline{f}, \text{T } p) = F) \in D \\
+    \quad \sigma, i \models_D B[ \overline{f} \mapsto \overline{F}, \overline{p} \mapsto\ evaluate(\overline{P}, \sigma(i))] \\
+    \quad \text{where } (N(\underline{\overline{\text{Form}}}\ \overline{f}, \text{T } p) = B) \in D \\
     \text{otherwise, if } i = 0 \text{ or } i = |\sigma|+1 \text{ then:} \\
-    \quad \text{rule } N \text{ is defined as }\underline{\text{max}}\text{ in } D
+    \quad \text{rule } N \text{ is defined as }\underline{\text{max}}/\underline{\text{min}}\text{ in } D
     \end{cases}
 \end{align*}
-where $\overline{F}$ and $\overline{P}$ represent tuples of type $\overline{\text{Form}}$ and $\overline{T}$ respectively
+where $\overline{F}$ and $\overline{P}$ represent tuples of type $\overline{\text{Form}}$ and $\overline{T}$ respectively.
 
@@
 \end{align*}
-and the sequence $\sigma=\alpha_{1},\ldots ,\alpha_{n}$, Property will evaluate to \textit{true} 
+and the sequence $\sigma=\alpha_{1},\ldots ,\alpha_{n}$, $\mathtt{Property}$ will evaluate to \textit{true} 
 \begin{align*}

### 2023-09-20 Paper/paper.tex
update

@@
 \end{align*}
-and the bold $\mathbf{Property(F)}$ will be evaluated as false, since it is the last of a big disjunction.
+and $\mathbf{Property(F)}$ will be evaluated as false, since it is the last of a big disjunction.
 \subsection{Evaluation algorithm}\label{algo}
@@
 These illustrations demonstrate the expressive power of EAGLE formalism as well as its numerous features.
-In all the following examples, the rule Always is the one defined in Section~\ref{sec:ltl}.
+In all the following examples, the rule $\mathtt{Always}$ is the one defined in Section~\ref{sec:ltl}.
 
@@
 as we can see, the \textit{type} field of a record schema describes the type of network packet, while the \textit{ip} field is the destination of the packet.
-The $\mathtt{SmurfAttack}$ rule determines whether the packet is a ping packet and whether the source's IP address is a broadcast ip, i.e.
+The $\mathtt{SmurfAttack}$ rule determines whether the packet is a ping packet and if the destination's IP address is a broadcast ip, i.e.
 all the host bits are set to 1.

### 2023-09-20 Presentation/presentation.tex
update

@@
     \begin{block}{Note.}    
-        The difference between maximal and minimal interpretation becomes important only when we are evaluating the at the boundaries of a trace; $\underline{\text{max}}$ rules evaluates at $\underline{\text{true}}$ at initial and final istants, while $\underline{\text{min}}$ rules evaluates at $\underline{\text{false}}$.
+        The difference between maximal and minimal interpretation becomes important only when we are evaluating the at the boundaries of a trace: $\underline{\text{max}}$ rules evaluates at $\underline{\text{true}}$ at initial and final istants, while $\underline{\text{min}}$ rules evaluates at $\underline{\text{false}}$.
     \end{block}    
@@
 
-    \includegraphics[scale=0.22]{images/semantics.png}
+    \includegraphics[scale=0.27]{images/semantics.png}
 \end{frame}
@@
     \end{align*}
-    where \textit{isBroadcast}() checks whether the source’s IP address is a broadcast ip, i.e. all the host bits are set to 1.
+    where \textit{isBroadcast}() checks whether the destination’s IP address is a broadcast ip, i.e. all the host bits are set to 1.
 \end{frame}
